# Analysis

The analysis phase is dedicated to the development, refinement, and comprehensive evaluation of the predictive models, meticulously designed to directly address the previously defined research questions.

## Methods

The modeling process is structured to address the two key research questions:

1.  identifying the most significant lifestyle and behavioral factors contributing to obesity in Mexico, Peru, and Colombia;

2.  predicting whether a person will be obese based on some given combinations of factors.

### Logistic Regression Model

To accurately address the key research questions, a logistic regression model will be employed to estimate the probability of individuals belonging to a category: *obese* or *not obese*. Weight and height will be excluded as predictors in the model because they are directly used to calculate BMI, which serves as the basis for the obesity levels categorized in the dataset. Including these variables would create a dependency between the predictors and the target variable, potentially biasing the analysis. By excluding weight and height, the focus shifts to behavioral and lifestyle factors, such as dietary habits, physical activity, and demographic characteristics, to better understand their influence on obesity risk.

While logistic regression provides a clear and interpretable framework for estimating probabilities, it inherently limits the analysis to a binary classification. This restriction prevents the exploration of the full spectrum of obesity levels, such as Obesity Type I, II, or III, as classified in the dataset. Despite this limitation, logistic regression is a robust method for quantifying the relationships between independent variables and the binary outcome. Feature selection techniques will ensure that only the most relevant predictors are retained, and the model's performance will be rigorously evaluated using metrics such as accuracy, precision, recall, F1-score, and ROC-AUC, ensuring reliable and actionable insights.

### Insights and Limitations

Regression analysis helps us understand how predictors influence outcomes, with logistic regression classifying individuals as obese or not obese. As already discussed in the previous sections, the dataset offers a mix of advantages and challenges: synthetic data ensures balanced representation but lacks the complexity of real-world patterns, while user-collected data adds variability but is prone to biases. Logistic regression simplifies the analysis by focusing on binary outcomes, leaving out the nuanced gradations of obesity, and assumes linearity, which may not fully capture complex relationships. Despite these limitations, the model offers insights into obesity risk, serving as a valuable exercise and foundation for future data explorations, even if not directly applicable to real-world scenarios.

## Objectives of the Selected Method

### Logistic Regression Model Development

**Data Loading and Processing**

To align with the requirements of a logistic regression model, it was necessary to modify the dataset's target variable. The original variable, obesity level, was a multi-class categorical variable representing varying degrees of obesity and non-obesity. Since logistic regression is designed for binary classification, the target variable was converted into a binary format. Individuals with a BMI ≥ 30 were classified as obese (1), while others were classified as non-obese (0). This transformation ensured compatibility with the logistic regression framework. Following this adjustment, the target variable was converted into a factor, and the dataset was reviewed for consistency and readiness for analysis.

```{r}
dataset <- read.csv(here("data/processed/dataset.csv"))
dataset$BMI <- dataset$weight / (dataset$height^2)
head(dataset$BMI) %>%
  tibble::enframe(name = "Row", value = "BMI") %>%
  kable(format = "html", caption = "First 6 BMI Values") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
  row_spec(0, bold = TRUE, background = "#f5f5f5")
summary_BMI <- summary(dataset$BMI)
summary_BMI %>%
  tibble::enframe(name = "Statistic", value = "Value") %>%
  kable(format = "html", caption = "Summary Statistics for BMI") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
  row_spec(0, bold = TRUE, background = "#f5f5f5")
dataset$Obesity <- ifelse(dataset$BMI >= 30, 1, 0)
dataset$Obesity <- as.factor(dataset$Obesity)
table(dataset$Obesity) %>%
  as.data.frame() %>%
  kable(format = "html", caption = "Frequency: obesity categories") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
  row_spec(0, bold = TRUE, background = "#f5f5f5")
head(dataset) %>%
  kable(format = "html", caption = "First 6 Rows of Updated Dataset") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#f5f5f5") %>%
  scroll_box(width = "100%", height = "400px")
```

Twelve predictors associated with obesity-related behaviors, dietary habits, physical activity, and lifestyle factors were selected for analysis. These variables, along with the binary target variable Obesity (1 = obese, 0 = not obese), formed the dataset for logistic regression modeling. The dataset was reviewed to ensure correct structure and readiness for analysis.

```{r}
predictors <- c("family_hist", "caloric_food", "vegetable_food", "nb_meal_day", "food_btw_meals", "smoke", "ch2o", "calorie_check", "physical_act", "use_tech", "freq_alcohol", "m_trans")
model_data <- dataset[, c("Obesity", predictors)]

str_output <- capture.output(str(model_data))
str_table <- data.frame(Structure = str_output, stringsAsFactors = FALSE)

str_table %>%
  kable(format = "html", caption = "Structure of Model Data") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#f5f5f5") %>%
  scroll_box(width = "100%", height = "400px")

head(model_data) %>%
  kable(format = "html", caption = "First 6 Rows of Model Data") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#f5f5f5") %>%
  scroll_box(width = "100%", height = "400px")
```

**Model Development**

Three regression models were employed to ensure a systematic and robust approach to predictor selection and model development. The null model, containing only the intercept, served as a baseline to represent predictions without the influence of any predictors. This provided a reference point to evaluate how much additional explanatory power was gained by including predictors.

The full model, incorporating all predictors, represented the maximum complexity allowable within the dataset. This model helped understand the potential contribution of each variable but carried the risk of overfitting due to its complexity.

The stepwise model, guided by the Akaike Information Criterion (AIC), balanced the simplicity and performance of the model. By iteratively evaluating the inclusion or exclusion of predictors, the stepwise procedure identified the subset of variables that significantly contributed to explaining the outcome while minimizing unnecessary complexity. This process ensured that the final model retained only the most relevant predictors, achieving optimal fit and generalizability. Using these three models allowed for a thorough comparison and the development of a parsimonious and effective predictive model.

```{r}
full_model <- glm(Obesity ~ ., data = model_data, family = binomial)
null_model <- glm(Obesity ~ 1, data = model_data, family = binomial)
stepwise_model <- step(null_model, scope = list(lower = null_model, upper = full_model), direction = "both", trace = FALSE)
```

Presented below is a comprehensive overview of the logistic regression models.

*Null Model*

```{r}
null_model_summary <- summary(null_model)

null_model_coef_table <- coef(null_model_summary) %>%
  as.data.frame() %>%
  tibble::rownames_to_column("Predictor")

null_model_coef_table %>%
  kable(format = "html", caption = "Coefficients of the Null Logistic Regression Model") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#f5f5f5") %>%
  scroll_box(width = "100%", height = "400px")
```

*Full Model*

```{r}
coef_table <- coef(summary(full_model)) %>%
  as.data.frame() %>%
  tibble::rownames_to_column("Predictor")

coef_table %>%
  kable(format = "html", caption = "Coefficients of the Full Logistic Regression Model") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#f5f5f5") %>%
  scroll_box(width = "100%", height = "400px")
```

*Stepwise Model*

```{r}
stepwise_summary <- summary(stepwise_model)

stepwise_coef_table <- coef(stepwise_summary) %>%
  as.data.frame() %>%
  tibble::rownames_to_column("Predictor")

stepwise_coef_table %>%
  kable(format = "html", caption = "Coefficients of the Stepwise Logistic Regression Model") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#f5f5f5") %>%
  scroll_box(width = "100%", height = "400px")
```

**Evaluation**

To evaluate the stepwise-selected model, predicted probabilities of obesity were generated for all individuals. These probabilities were converted into binary classifications using a threshold of 0.5. A confusion matrix was constructed to assess the model's performance, providing key metrics such as accuracy, sensitivity, specificity, precision, and F1-score.

```{r}
library(caret)
library(pROC)
# Predict probabilities from the stepwise model
predicted_probs <- predict(stepwise_model, type = "response")

# Convert probabilities to binary classes (using a threshold of 0.5)
predicted_classes <- ifelse(predicted_probs >= 0.5, 1, 0)

# Create a confusion matrix (comparison of predicted vs. actual values)
conf_matrix <- confusionMatrix(as.factor(predicted_classes), model_data$Obesity)

conf_matrix_table <- as.data.frame(conf_matrix$table)
colnames(conf_matrix_table) <- c("Actual", "Predicted", "Count")
conf_matrix_table <- conf_matrix_table %>%
  group_by(Actual) %>%
  mutate(Percentage = round((Count / sum(Count)) * 100, 2)) %>%
  ungroup()

conf_matrix_table %>%
  kable(format = "html", caption = "Confusion Matrix: Predicted vs Actual Values with Percentages") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#f5f5f5") %>%
  column_spec(1, bold = TRUE, width = "150px") %>%
  column_spec(2, width = "150px") %>%
  column_spec(3, width = "100px") %>%
  column_spec(4, width = "100px") %>%
  scroll_box(width = "100%", height = "400px")

performance_metrics_vertical <- as.data.frame(conf_matrix$overall) %>%
  tibble::rownames_to_column("Metric") %>%
  tidyr::pivot_longer(cols = -Metric, names_to = NULL, values_to = "Value")
performance_metrics_vertical %>%
  kable(format = "html", caption = "Confusion Matrix Overall Performance Metrics") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#f5f5f5") %>%
  scroll_box(width = "100%", height = "400px")

class_metrics_vertical <- as.data.frame(conf_matrix$byClass) %>%
  tibble::rownames_to_column("Metric") %>%
  tidyr::pivot_longer(cols = -Metric, names_to = NULL, values_to = "Value")
class_metrics_vertical %>%
  kable(format = "html", caption = "Class-Level Performance Metrics") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#f5f5f5") %>%
  scroll_box(width = "100%", height = "400px")
```

Additionally, the ROC curve and AUC were calculated to further evaluate the model’s discriminative ability. The ROC curve visualizes the trade-off between sensitivity and specificity, while the AUC quantifies the model's ability to distinguish between obese and non-obese individuals.

```{r, message=FALSE}
# Compute ROC curve using the actual class labels and predicted probabilities
roc_curve <- pROC::roc(model_data$Obesity, predicted_probs)
auc_value <- pROC::auc(roc_curve)
print(auc_value)
```

**Results Visualization**

To assess the distribution of predicted probabilities, a scatter plot was created with observations color-coded by their actual class. This visualization provides a clear overview of the model's predictions and potential misclassifications.

```{r}
plot(predicted_probs, col = ifelse(model_data$Obesity == 1, "blue", "red"), pch = 16, xlab = "n° Observation", ylab = "Predicted Probability", main = "Predicted Probabilities of Obesity", cex = 0.6)
legend("bottomright", legend = c("Obese", "Not Obese"), col = c("blue", "red"), pch = 16)
```

For additional clarity, the ROC curve was plotted to visually represent the model's performance.

```{r}
plot(roc_curve, col = "blue", main = "ROC Curve", lwd = 3, xlim = c(0, 1), ylim = c(0, 1.05), xlab = "False Positive Rate", ylab = "True Positive Rate", cex.main = 1.5, cex.lab = 1.2, cex.axis = 1.1)
legend("topright", legend = paste("AUC =", round(auc_value, 3)), lwd = 0, cex = 1.2, bty = "n")
grid()
```

**Predicting Obesity Probability**

To test the model’s ability to predict the probability of individuals becoming obese, six distinct profiles were created, representing a diverse range of lifestyles. Each profile was carefully designed to highlight specific behavioral, dietary, and lifestyle patterns.

The first individual represents a high-risk case for obesity. This person has a family history of being overweight, frequently consumes high-calorie foods and snacks, and eats very few vegetables. They have five meals a day, drink only 0.5 liters of water daily, and do no physical activity. Additionally, they spend 10 hours a day using technology, consume alcohol consistently, and rely primarily on public transportation for mobility.

The second individual exemplifies a very healthy lifestyle. They have no family history of being overweight, rarely consume high-calorie foods or snacks, and eat a large amount of vegetables. Their diet consists of very few meals per day, complemented by a high water intake of 4 liters daily. They do not monitor calorie intake but engage in physical activity five times a week. They walk as their primary mode of transportation, do not consume alcohol, and spend only 0.5 hours daily using technology.

The third individual exhibits a balanced lifestyle but shows some risk factors. This person has a family history of being overweight, frequently consumes snacks and high-calorie foods, and eats a moderate amount of vegetables. They have three meals a day, drink 1 liter of water, and monitor their calorie intake. However, they engage in physical activity only once a week, use technology for 8 hours daily, use motorbike as transportation vehicle, and occasionally consume alcohol.

The fourth individual is physically active and health-conscious. They have no family history of being overweight, do not frequently consume high-calorie foods, but snack occasionally. They eat a lot of vegetables, have a small number of meals per day, and drink 3 liters of water daily. They do not monitor calorie intake but exercise three times a week and use a bicycle for transportation. They consume alcohol frequently but spend only an hour daily using technology.

The fifth individual represents another high-risk case due to a sedentary lifestyle. They have a family history of being overweight, frequently consume high-calorie foods and snacks, and eat very few vegetables. They have four meals a day, drink 2 liters of water, and do no physical activity. They spend 6 hours daily using technology, consume alcohol moderately, and rely on public transportation.

The sixth individual leads a very active lifestyle but has some risk factors due to alcohol and transportation choices. They have no family history of being overweight, do not frequently consume high-calorie foods or snacks, and eat a large amount of vegetables. They have two meals per day, drink 1.5 liters of water, and do not monitor calorie intake. However, they engage in physical activity four times a week, use a motorbike for transportation, do not consume alcohol, and spend 2 hours daily using technology.

These six profiles were designed to test the model’s capacity to handle a wide variety of real-world scenarios, ensuring it can effectively predict obesity probabilities across diverse populations.

```{r}
new_data <- data.frame(
  family_hist = factor(c("yes", "no", "yes", "no", "yes", "no"), 
                                          levels = c("yes", "no")),
  caloric_food = factor(c("yes", "no", "yes", "no", "yes", "no"), 
                levels = c("yes", "no")),
  vegetable_food = c(1, 5, 2, 4, 1, 3),
  nb_meal_day = c(5, 1, 3, 2, 4, 2),
  food_btw_meals = factor(c("Frequently", "Sometimes", "Always", "Sometimes", "Frequently", "Always"), 
                levels = c("Frequently", "Sometimes", "Always")),
  smoke = factor(c("no", "yes", "no", "yes", "yes", "no"), 
                 levels = c("yes", "no")),
  ch2o = c(0.5, 4, 1, 3, 2, 1.5),
  calorie_check = factor(c("yes", "no", "yes", "no", "yes", "no"), 
               levels = c("yes", "no")),
  physical_act = c(0, 5, 1, 3, 0, 4),
  use_tech = c(10, 0.5, 8, 1, 6, 2),
  freq_alcohol = factor(c("Always", "Never", "Sometimes", "Always", "Sometimes", "Never"), 
                levels = c("Sometimes", "Frequently", "Always", "Never")),
  m_trans = factor(c("Public_Transportation", "Walking", "Motorbike", "Bike", "Public_Transportation", "Motorbike"), 
                  levels = c("Public_Transportation", "Walking", "Bike", "Motorbike"))
)

predicted_probs_new <- predict(stepwise_model, new_data, type = "response")

probability_table <- tibble(
  Reference = 1:length(predicted_probs_new),
  Predicted_Probability = predicted_probs_new,
  Probability_Percentage = predicted_probs_new * 100
)

probability_table %>%
  kbl(format = "html", caption = "Predicted Probabilities") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
  column_spec(1, width = "100px") %>%
  scroll_box(width = "100%", height = "400px")
```

## Results

**Null Model**

Under the null model, only the intercept was included to predict the probability of obesity, without accounting for any behavioral or lifestyle predictors. The estimated intercept value of -0.133403 suggests that, in the absence of explanatory variables, the average probability of obesity in the analyzed population is below 50%. This indicates that obesity is not the default outcome when no additional factors are considered. The statistical significance of the intercept (p-value = 0.0024) reinforces the reliability of this baseline estimate, demonstrating that the model is not predicting obesity purely by chance. However, this simplistic model serves as a benchmark against which more complex models and does not incorporate any individual or contextual factors that are hypothesized to influence obesity outcomes.

**Full Logistic Regression Model**

The results of the full logistic regression model highlight significant statistical relationships between key variables and the likelihood of obesity. Among the predictors, family history of being overweight demonstrated the strongest association (coefficient = 3.697, p \< 0.001), emphasizing the critical role of genetic predisposition in influencing obesity risk.

Dietary habits were also prominently associated with obesity. Frequent consumption of high-calorie foods (FAVC) showed a strong positive effect (coefficient = 2.071, p \< 0.001), reinforcing the role of energy-dense diets in driving weight gain. Interestingly, vegetable consumption (FCVC) was also positively associated (coefficient = 0.880, p \< 0.001). While unexpected, this may reflect broader dietary patterns, where higher vegetable intake coexists with other risk behaviors in certain populations.

Snacking behavior revealed complex relationships. Frequent eating between meals (CAECFrequently) was negatively associated with obesity (coefficient = -2.064, p = 0.0005), suggesting that regular snacking might displace larger, energy-dense meals or align with healthier dietary practices. Conversely, occasional snacking (CAECSometimes) increased obesity risk (coefficient = 1.269, p = 0.0057), which may reflect irregular or opportunistic consumption of unhealthy snacks.

Behavioral and lifestyle factors further enriched the analysis. Smoking (SMOKE) was positively associated with obesity (coefficient = 1.055, p = 0.022), potentially indicating correlations with sedentary behaviors or poor dietary choices often linked to smoking habits. Transportation mode had contrasting effects: reliance on public transportation (MTRANS: Public_Transportation) was positively associated (coefficient = 0.570, p \< 0.001), while walking as a primary mode (MTRANS: Walking) was strongly protective (coefficient = -1.903, p = 0.0035). These findings underscore the importance of physical activity embedded in daily routines as a key determinant of obesity risk.

A threshold of **p \< 0.05** was used to determine statistical significance for all predictors, with highly significant variables (e.g., p \< 0.001) providing robust evidence of their association with obesity. Although some predictors, such as water intake (CH2O) and physical activity frequency (FAF), showed weaker associations, their coefficients suggest that they may still contribute to obesity risk in specific contexts. Overall, the model effectively captured the multifaceted nature of obesity, highlighting the interplay between genetic predisposition, dietary behaviors, and lifestyle choices. These results provide a comprehensive foundation for refining predictive models and developing targeted interventions to mitigate obesity in diverse populations.

**Stepwise Logistic Regression Model**

The stepwise logistic regression model was designed to isolate the most influential predictors of obesity while maintaining a balance between model complexity and interpretability. By employing the Akaike Information Criterion (AIC) as a criterion for iterative inclusion and exclusion of variables, the model retained only those predictors that significantly contributed to explaining variations in obesity outcomes. This process resulted in a parsimonious and focused model centered on key behavioral, dietary, and lifestyle factors.

Among the retained predictors, family history of overweight emerged as the most significant variable, with a coefficient of 3.679 (p \< 0.001). This finding underscores the critical influence of genetic predisposition on obesity, reinforcing the notion that familial trends play a central role in determining obesity risk. Similarly, frequent consumption of high-calorie foods was strongly linked to obesity, with a coefficient of 2.117 (p \< 0.001). This emphasizes the fundamental role of dietary patterns, particularly the consumption of energy-dense foods, in driving obesity outcomes.

The model also uncovered complex associations between meal timing and obesity. Frequent snacking between meals was found to have a significant negative association with obesity (coefficient = -2.024, p = 0.0006), suggesting that certain structured eating behaviors may mitigate obesity risk. Conversely, occasional snacking exhibited a positive association with obesity (coefficient = 1.361, p = 0.0028). These findings indicate that both the frequency and context of eating outside traditional meals may play a nuanced role in weight regulation.

Modes of transportation were also identified as significant predictors. Use of public transportation was positively associated with obesity (coefficient = 0.602, p \< 0.001), potentially reflecting a sedentary component in daily commuting. In contrast, walking as a primary mode of transportation was negatively associated with obesity (coefficient = -1.853, p = 0.0044), highlighting the protective role of physical activity integrated into daily routines.

Additional predictors included smoking, which was positively correlated with obesity (coefficient = 1.021, p = 0.024), and calorie monitoring, which exhibited a strong protective effect (coefficient = -2.673, p \< 0.001). Both physical activity (coefficient = -0.347, p \< 0.001) and time spent using technology (coefficient = -0.417, p \< 0.001) were negatively associated with obesity, reinforcing the importance of active lifestyles in maintaining healthy weight.

This stepwise approach distilled the most impactful predictors of obesity from a broader set of variables, providing a refined understanding of the key drivers of obesity in the studied populations. By focusing on a subset of significant predictors, the model not only enhances interpretability but also ensures practical utility in guiding targeted interventions and public health policies. These findings represent a critical step toward addressing the multifactorial nature of obesity with evidence-based strategies.

**Model Evaluation**

The evaluation of the stepwise logistic regression model demonstrates its effectiveness in predicting obesity based on a set of behavioral and lifestyle predictors. The overall accuracy of the model, calculated at **75.89%** with a 95% confidence interval between **74.04%** and **77.72%,** highlights its reliability in distinguishing obese and non-obese individuals. The kappa statistic of **0.5227** indicates moderate agreement between predicted and actual classifications, confirming that the model's predictions are well above random chance. Both the accuracy and McNemar test p-values (\<0.001) further validate the statistical robustness of the model.

From a class-level perspective, the sensitivity, or the model’s ability to correctly classify obese individuals, was **67.30%.** This suggests that while the model captures the majority of true positive cases, a proportion of obese individuals are still misclassified as non-obese. In contrast, the specificity was higher, at **85.73%,** indicating a strong ability to correctly identify non-obese individuals. Precision, which quantifies the proportion of correctly identified obese individuals among those predicted as obese, reached **84.35%,** highlighting the model's reliability in minimizing false positive classifications. The F1 score of **74.86%** demonstrates a solid balance between precision and recall, further emphasizing the model's overall efficacy. Balanced accuracy, which averages sensitivity and specificity, was **76.51%,** underscoring the model's consistent performance across both classes.

The receiver operating characteristic (ROC) curve provides additional validation of the model's predictive capacity. The area under the curve (AUC) was calculated as **0.856**, which reflects a high level of discriminative power. This indicates that the model is highly capable of distinguishing between obese and non-obese individuals, with a strong ability to balance false positive and false negative rates across various decision thresholds.

Taken together, these results confirm the predictive reliability and statistical soundness of the stepwise logistic regression model. The findings suggest that the model can serve as a valuable tool for identifying key risk factors associated with obesity and for making predictions in diverse populations. While the model demonstrates strong performance metrics overall, particularly in terms of specificity and precision, future iterations may aim to enhance sensitivity to minimize the misclassification of obese individuals. These insights provide a strong foundation for the application of this model in public health contexts and targeted interventions.

**Performance**

The stepwise logistic regression model was applied to six hypothetical profiles representing diverse behavioral and lifestyle patterns, yielding predicted probabilities ranging from near zero to slightly above 3%. This distribution reflects the model’s conservative thresholds, which prioritize specificity and minimize false positives. While this ensures reliable identification of individuals at high risk, it may underestimate the probabilities for those with moderate or mixed risk factors. The scatter plot of predicted probabilities confirms the model’s ability to differentiate obese from non-obese individuals, particularly at the extremes of the probability range. However, the overlap observed in mid-range probabilities indicates challenges in confidently classifying borderline cases.

A notable strength of the model is its interpretability and its reliance on behavioral and lifestyle predictors, which are directly actionable in public health contexts. Its high specificity makes it particularly well-suited for targeted interventions, reducing the risk of over-diagnosis and unnecessary allocation of resources. However, the conservative nature of the model comes at the cost of potentially overlooking individuals at intermediate risk levels. Additionally, while the selected predictors are practical and relevant, the exclusion of psychosocial and environmental factors—such as socioeconomic status or urban versus rural living conditions—limits the model's applicability in settings where these variables are significant contributors to obesity.

Overall, the stepwise logistic regression model provides a reliable and practical framework for predicting obesity and guiding public health strategies. Its strengths in specificity and precision are clear, but future improvements should focus on enhancing sensitivity and expanding the range of predictors to improve its performance across diverse populations and more complex real-world scenarios. These refinements will ensure the model’s continued relevance and utility in addressing the multifaceted nature of obesity.
