# Data

## Sources

The dataset utilized in this project was obtained from the UCI Machine Learning Repository, a reputable and extensively used platform for data science and machine learning projects. Originally compiled by researchers at the Universidad de la Costa, Colombia, the dataset combines 77% synthetically generated data with 23% real-world data collected through a structured online survey. The synthetic data, created using the Synthetic Minority Over-sampling Technique (SMOTE) in Weka, addresses class imbalance, enhancing the dataset's suitability for machine learning tasks. The real-world data, gathered from 498 participants over a 30-day period, captures detailed self-reported information on dietary habits, physical activity levels, and demographic characteristics. While synthetic data introduces uniformity and balance, it inherently lacks the complexity of real-world variability, and the user-collected data, though authentic, is susceptible to self-reporting biases and sampling limitations. These characteristics, along with the dataset’s diverse origins, make it an invaluable resource for simulating real-world challenges in healthcare analytics.

## Description

The dataset consists of 2111 records and 17 attributes, offering a detailed examination of the factors contributing to obesity. The attributes represent a mix of categorical and continuous variables, providing insights into demographic, lifestyle, and behavioral factors. In greater detail, an interactive table was designed to provide a comprehensive summary of the dataset's variables.

```{r}
library(here)
library(knitr)
# Main features of the dataset
dataset_raw <- read.csv(here("data/raw/dataset_raw.csv"))
head(dataset_raw) %>%
  kbl(format = "html", caption = "First 6 Rows of the Raw Dataset") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#f0f0f0") %>%
  scroll_box(width = "100%", height = "400px")

val_meaning <- c("Indicates the gender of the individual (Male/Female).", "Represents the age of participants in years.", "The height of individuals in meters.", "The weight of participants in kilograms.", "Indicates whether a family member has suffered from overweight (Yes/No).", "Indicates if participants frequently consume high-caloric foods (Yes/No).", "Scaled from 1 to 3, reflects how often vegetables are consumed (1 = Never, 3 = Always).", "Indicates the typical number of main meals consumed daily.", "Describes how often participants eat between meals (e.g., No, Sometimes, Frequently, Always).", "Indicates whether participants smoke (Yes/No).", "Scaled from 1 to 3, reflecting daily water intake (1 = Less than 1 liter, 3 = More than 2 liters).", "Whether participants monitor their calorie intake (Yes/No).", "Scaled from 0 to 4, indicating days of physical activity per week (0 = None, 4 = 4-5 days).", "Reflects daily time spent on technological devices, in hours.", "Indicates the frequency of alcohol consumption (e.g., I don't drink, Sometimes, Frequently, Always).", "Describes the primary mode of transportation (e.g., Walking, Public Transportation, Automobile).", "The target variable, classifying obesity levels into categories such as Normal Weight, Overweight (Levels I and II), and Obesity (Types I, II, III).")
desc_table <- tibble::tibble(Name = colnames(dataset_raw), Type = sapply(dataset_raw, class), Meaning = val_meaning)
desc_table %>%
  kbl(format = "html", caption = "Variable Descriptions") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#f0f0f0") %>%
  column_spec(1, bold = TRUE, width = "200px") %>%
  column_spec(2, width = "150px") %>%
  column_spec(3, width = "500px") %>%
  scroll_box(width = "100%", height = "500px")
```

The dataset underwent a thorough preprocessing phase, including normalization of continuous variables, encoding of categorical data, and removal of missing or atypical entries to ensure high-quality analysis. Class imbalance was addressed using the SMOTE (Synthetic Minority Oversampling Technique), generating synthetic data while carefully avoiding noise or artificial patterns. The final dataset comprises 77% synthetic data, which enhances balance and diversity, and 23% real-world data, adding authenticity. This combination allows for a comprehensive analysis of obesity-related factors, while recognizing potential biases, such as inaccuracies in self-reported information.

## Wrangling

Essential libraries for data manipulation, visualization, and clustering are loaded to begin the wrangling process and support subsequent analysis. Each package is utilized for its specific functionality, facilitating efficient and streamlined analysis:

-   dplyr: for data manipulation (e.g., filtering, summarizing);

-   tidyr: for data tidying (e.g., reshaping);

-   ggplot2: for visualization;

-   corrplot: for correlation matrix visualization;

-   ggridges: for creating ridge plots;

-   cluster: for clustering algorithms;

-   reshape2: for data reshaping, especially during visualization.

```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(corrplot)
library(ggridges)
library(cluster)
library(reshape2)
```

Column names are renamed to enhance clarity and improve usability during the analysis. The updated names are designed to be shorter and more intuitive, ensuring ease of reference while retaining their original meaning and context. This adjustment simplifies code readability and helps streamline data manipulation tasks, particularly in complex analytical workflows.

```{r}
  dataset <- dataset_raw %>%
  rename(
    family_hist = family_history_with_overweight,
    obesity_lev = NObeyesdad,
    caloric_food = FAVC,
    vegetable_food = FCVC,
    nb_meal_day = NCP,
    food_btw_meals = CAEC,
    ch2o = CH2O,
    smoke = SMOKE,
    calorie_check = SCC,
    physical_act = FAF,
    freq_alcohol = CALC,
    use_tech = TUE,
    m_trans = MTRANS,
    gender = Gender,
    age = Age,
    weight = Weight,
    height = Height
  )
```

The structure of the dataset is examined to identify the data types of each variable, providing critical insights for subsequent data preparation. Understanding the data types helps pinpoint columns requiring transformations, such as converting categorical variables to factors or standardizing numeric variables for analysis.

```{r}
str_output <- capture.output(str(dataset))
str_table <- data.frame(Structure = str_output, stringsAsFactors = FALSE)
str_table %>%
  kbl(format = "html", caption = "Original structure of the Dataset") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#f0f0f0") %>%
  scroll_box(width = "100%", height = "400px")


dataset <- dataset %>%
  mutate(
    gender = as.factor(gender),
    family_hist = as.factor(family_hist),
    caloric_food = as.factor(caloric_food),
    smoke = as.factor(smoke),
    calorie_check = as.factor(calorie_check),
    m_trans = as.factor(m_trans),
    obesity_lev = factor(obesity_lev, 
                         levels = c("Insufficient_Weight", "Normal_Weight", 
                                    "Overweight_Level_I", "Overweight_Level_II", 
                                    "Obesity_Type_I", "Obesity_Type_II", "Obesity_Type_III"), 
                         ordered = TRUE),
    food_btw_meals = factor(ifelse(food_btw_meals == "no", "No", food_btw_meals), 
                            levels = c("No", "Sometimes", "Frequently", "Always"), 
                            ordered = TRUE),
    freq_alcohol = factor(ifelse(freq_alcohol == "no", "No", freq_alcohol), 
                          levels = c("No", "Sometimes", "Frequently", "Always"), 
                          ordered = TRUE))


str_output <- capture.output(str(dataset))
str_table <- data.frame(Structure = str_output, stringsAsFactors = FALSE)
str_table %>%
  kbl(format = "html", caption = "Dataset Structure") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#f5f5f5") %>%
  scroll_box(width = "100%", height = "400px")
```

The transformations ensured the dataset was ready for analysis by restructuring categorical and ordinal variables to meet modeling requirements. Converting categorical variables into factors standardized their representation, reducing ambiguity and improving compatibility with statistical models. For ordinal variables, levels were explicitly ordered to preserve their logical progression and enhance interpretability, allowing for meaningful comparisons across categories.

The updated structure was reviewed to confirm the accuracy of these adjustments, providing confidence in the preprocessing steps. While further transformations like normalization were not applied, the focus on categorical and ordinal adjustments established a strong foundation for reliable and interpretable analysis. In particular, the levels of obesity categories, food consumption between meals, and frequency of alcohol use were arranged to reflect increasing severity or frequency, ensuring these variables captured their intended relationships and supported clear, accurate insights into the data.

Now, a numerical version of the dataset, called "dataset_num", is created by transforming categorical variables into numerical values, ensuring compatibility with statistical analyses while maintaining logical relationships and interpretability. This numerical transformation is specifically essential for developing the correlation matrix, as it requires all variables to be in numeric format to analyze their relationships effectively.

The presence of potential missing values in the transformed dataset is checked and visualized to confirm data integrity and ensure no issues have been introduced during the conversion process.

```{r}

dataset_num <- dataset %>%
  mutate(obesity_lev = recode(obesity_lev,
                              "Insufficient_Weight"=1,
                              "Normal_Weight" = 2,
                              "Overweight_Level_I" = 3,
                              "Overweight_Level_II" = 4,
                              "Obesity_Type_I" = 5,
                              "Obesity_Type_II" = 6,
                              "Obesity_Type_III" = 7,
  ))

dataset_num <- dataset %>%
  mutate(freq_alcohol = recode(freq_alcohol,
                               "No"=1,        
                               "Sometimes"=2, 
                               "Frequently" =3,
                               "Always"  =4 
  ))

dataset_num <- dataset %>%
  mutate(m_trans = recode(m_trans,
                          "Automobile"=1,
                          "Bike"=2,
                          "Motorbike"=3,
                          "Public_Transportation"=4,
                          "Walking"=5,
  ))

dataset_num <- dataset %>%
  mutate(food_btw_meals = recode(food_btw_meals,
                                 "No"=0,
                                 "Sometimes"=1 ,
                                 "Frequently"=2,
                                 "Always"=3
  )
  )

dataset_num <- dataset_num%>%
  mutate(calorie_check = recode(calorie_check,
                                "no"=0,
                                "yes"=1 ,
  ))

dataset_num <- dataset_num %>%
  mutate(across(where(is.factor), ~ as.numeric(.)))


str_output <- capture.output(str(dataset_num))
table_num_str <- data.frame(Structure = str_output, stringsAsFactors = FALSE)

table_num_str %>%
  kbl(format = "html", caption = "Structure of the Numerical Dataset") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#f5f5f5") %>%
  scroll_box(width = "100%", height = "400px")


nb_na<- colSums(is.na(dataset_num))
nb_na %>%
  kbl(format = "html", caption = "Presence of Potential NA Values in the Dataset") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"), 
    full_width = FALSE, 
    position = "left"
  ) %>%
  column_spec(1, width = "100px") %>%
  column_spec(2, width = "80px") %>%
  row_spec(0, bold = TRUE, background = "#f5f5f5") %>%
  scroll_box(width = "100%", height = "400px")
```

The test results confirmed the absence of any NA values in the dataset, indicating that all variables were successfully converted to numeric format without compromising data integrity.

## Spotting Mistakes and Missing Data

**Check for missing values**

To ensure data integrity, missing values in the dataset are examined by counting "NA" values in each column, providing a clear view of dataset completeness. The results are presented in a formatted table for easy interpretation, with styling applied for readability and a scrollable box to handle larger datasets. This process facilitates prompt handling of missing data through appropriate strategies.

```{r}
missing_values <- colSums(is.na(dataset))
missing_values %>%
  kbl(format = "html", caption = "Missing Values in Each Column") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width=FALSE, position = "center") %>%
  column_spec(1, width = "100px") %>%
  column_spec(2, width = "80px") %>%
  row_spec(0, bold = TRUE, background = "#f0f0f0") %>%
  scroll_box(width = "100%", height = "400px")
```

The analysis confirms that all columns contain complete data, with no missing values identified. This completeness ensures a robust foundation for subsequent analysis, eliminating the need for immediate data cleaning related to missing entries.

**Check for duplicates**

The dataset is examined for duplicated rows to ensure data integrity and eliminate redundancy. Identifying and addressing duplicates is a crucial step in data preprocessing, as redundant entries can skew analysis results and lead to misleading conclusions. This process involves systematically scanning the dataset for identical rows and quantifying their occurrence.

```{r}
duplicated_rows <- sum(duplicated(dataset))
duplicated_rows
```

The detection of 24 duplicated rows in the dataset highlights the need for further preprocessing to ensure data integrity, as these redundant entries could skew analysis if not properly handled.

```{r}
dataset <- dataset %>%
distinct()

nrow(dataset)
any(duplicated(dataset))
```

The dataset was refined by removing duplicate entries to ensure that only unique rows are retained. A verification step confirmed that no duplicates remain, ensuring the dataset's integrity and reliability for further analysis.

## Listing Anomalies and Outliers

A bar chart was created to visualize the distribution of obesity levels, providing a clear overview of class frequencies within the dataset. Particular attention is given to obesity levels, as this variable serves as the dependent variable in the predictive model to be developed later.

```{r}
g1 <- ggplot(dataset, aes(x = obesity_lev)) +
  geom_bar(fill = "skyblue", color = "black") +
  theme_minimal() +
  labs(
    title = "Class Distribution of Obesity Levels",
    x = "Obesity Level",
    y = "Count"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) #Adjusted the text for clarity

plotly_plot <- ggplotly(g1)
plotly_plot
```

The chart highlights a balanced distribution across all obesity levels, demonstrating the effectiveness of SMOTE in addressing class imbalance. By equalizing the representation of each category, the dataset becomes more reliable for analysis, reducing biases and ensuring a fair evaluation of patterns within the data. On the other hand, the synthetic data introduced by SMOTE may not fully reflect real-world variability, potentially leading to artificial patterns that could affect the interpretability of results.

A density plot was generated to visualize the age distribution across different obesity levels, providing insights into patterns and trends within the data.

```{r}
g2 <- ggplot(dataset, aes(x = age, fill = obesity_lev)) +
  geom_density(alpha = 0.5) +
  theme_minimal() +
    labs(
    title = "Age Distribution by Obesity Levels",
    x = "Age",
    y = "Density",
    fill = "Obesity Level")
plotly_plot1 <- ggplotly(g2)
plotly_plot1
```

This graph provides a detailed view of the age distribution across obesity levels and offers insight into the impact of the SMOTE algorithm in balancing the dataset. The distributions show distinct separation among obesity categories, with younger ages predominantly associated with lower obesity levels (e.g., Insufficient Weight and Normal Weight), while older ages are more prevalent in higher obesity categories (e.g., Obesity Type II and III).

Notably, sharp peaks in the density curves, such as the one around age 30 in "Obesity Type I," could indicate potential artifacts introduced during the synthetic data generation process. While these patterns align with logical demographic trends, they highlight the need for further validation to ensure that such separations and peaks represent realistic population characteristics rather than biases from data augmentation. Overall, the dataset reflects clear and interpretable patterns, but these observations suggest the importance of cautious interpretation and robust validation in subsequent analyses.

Summary statistics were computed for key variables across obesity levels to identify potential anomalies or patterns, providing a clearer understanding of how age, height, and weight vary within each category.

```{r}
dataset_stat <- dataset %>%
  group_by(obesity_lev) %>%
  summarize(
    Age_Mean = mean(age, na.rm = TRUE),
    Age_SD = sd(age, na.rm = TRUE),
    Height_Mean = mean(height, na.rm = TRUE),
    Height_SD = sd(height, na.rm = TRUE),
    Weight_Mean = mean(weight, na.rm = TRUE),
    Weight_SD = sd(weight, na.rm = TRUE)
  )
dataset_stat %>%
  kbl(format = "html", caption = "Summary Statistics by Obesity Level", digits = 1) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#f5f5f5") %>%
  scroll_box(width = "100%", height = "400px")
```

The summary statistics reveal distinct differences across obesity levels. As expected, weight increases progressively with higher obesity categories, accompanied by slightly larger variations in standard deviation. Interestingly, height remains relatively constant across categories, suggesting it plays a limited role in distinguishing obesity levels. The age distribution shows a notable shift, with younger individuals dominating the lower obesity levels and a broader age range in higher levels, highlighting potential demographic patterns worth further exploration. These insights confirm the logical trends in the dataset, providing confidence in its structure while emphasizing the need for further analysis of these relationships.

Clustering was performed using k-means to explore the dataset's structure and assess the coherence of the groups, with the silhouette score calculated to evaluate the quality and separation of the clusters.

```{r}
library(cluster)
set.seed(123)
kmeans_res <- kmeans(select(dataset, where(is.numeric)), centers = length(unique(dataset$obesity_lev)))
silhouette_score <- silhouette(kmeans_res$cluster, dist(select(dataset, where(is.numeric))))
mean_silhouette_score <- mean(silhouette_score[, "sil_width"])
mean_silhouette_score
```

The mean silhouette score of approximately 0.456 indicates moderate cohesion within clusters and reasonable separation between them. This suggests that the clusters, representing different obesity levels, are distinguishable but not excessively isolated. The result reflects a balance between natural class separability and the effects of data augmentation with SMOTE, which appears to have effectively balanced the dataset without introducing significant distortions. These findings provide confidence in the dataset’s suitability for clustering-based exploration while highlighting the importance of further validation to ensure the robustness of the observed patterns.

## Correlation Analysis

To explore relationships among variables and their association with obesity levels, a correlation matrix was computed. The analysis focuses on identifying the strength and direction of correlations between "obesity_lev" (the dependent variable) and other predictors, such as physical activity, frequency of alcohol consumption, and dietary habits. By ordering variables based on their correlation with "obesity_lev", the matrix highlights the most influential factors in determining obesity levels. A heatmap visualization was then created to provide an intuitive representation of these relationships, with a gradient color scale indicating the strength of positive and negative correlations. This approach facilitates the identification of key variables for further analysis and modeling.

```{r}
#Assuming dataset_num is already defined and contains the relevant columns
cor_matrix <- cor(dataset_num %>% select("physical_act", "freq_alcohol", "obesity_lev", "age", "weight","height", "family_hist", "caloric_food", "vegetable_food", "food_btw_meals", "use_tech", "ch2o", "m_trans", "smoke","nb_meal_day", "calorie_check", "gender"),use = "complete.obs")

#Extract the correlations with 'obesity_lev'
cor_with_obesity_lev <- cor_matrix["obesity_lev",]

#Order variables by their correlation with 'obesity_lev'
ordered_vars <- names(sort(cor_with_obesity_lev, decreasing = TRUE))

#Reorder the correlation matrix based on this order
cor_matrix_ordered <- cor_matrix[ordered_vars, ordered_vars]

#Melt the ordered correlation matrix into long format
cor_long <- melt(cor_matrix_ordered)

g3 <- ggplot(cor_long, aes(x = Var1, y = Var2, fill = value)) + 
    geom_tile() + 
    geom_text(aes(label = round(value, 2)), color = "black", size = 2.5, vjust = 0.5, hjust = 0.5) +
    scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
    labs(title = "Correlation Heatmap Ordered by Obesity Level", x = "Variables", y
       = "Variables") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        axis.text.y = element_text(angle = 45, vjust = 1))

plot3 <- ggplotly(g3)
plot3
# Create the heatmap with correlation values

# Assuming dataset_num is already defined and contains the relevant columns
cor_matrix <- cor(dataset_num %>% select("physical_act", "freq_alcohol", "obesity_lev", "age", "weight", "family_hist", "caloric_food", "vegetable_food", "food_btw_meals", "use_tech","ch2o", "height", "calorie_check", "gender"), use = "complete.obs")

# Extract the correlations with "obesity_lev"
cor_with_obesity_lev <- cor_matrix["obesity_lev",]

# Order variables by their correlation with 'obesity_lev'
ordered_vars <- names(sort(cor_with_obesity_lev, decreasing = TRUE))

# Reorder the correlation matrix based on this order
cor_matrix_ordered <- cor_matrix[ordered_vars, ordered_vars]

# Melt the ordered correlation matrix into long format
cor_long <- melt(cor_matrix_ordered)

g4 <- ggplot(cor_long, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() +
  geom_text(aes(label = round(value, 2)), color = "black", size = 2.5, vjust = 0.5
            , hjust = 0.5) + # Center text within tiles
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  labs(title = "Correlation Heatmap Ordered by Obesity Level", x = "Variables", y
       = "Variables") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        axis.text.y = element_text(angle = 45, vjust = 1) 
  )
plot4 <- ggplotly(g4)
plot4
```

The correlation matrices provide valuable insights into the relationships between variables and their association with obesity levels. As expected, weight exhibits a very strong positive correlation with obesity level, reinforcing its central role in defining the target variable. Family history of obesity and caloric food consumption also show moderate positive correlations, highlighting their relevance as predictive factors.

Conversely, variables such as physical activity and food consumption between meals exhibit weak or negative correlations, suggesting that their influence on obesity levels is less pronounced. These patterns align with logical trends but also underscore the need for careful consideration of multicollinearity and the relative importance of variables in predictive modeling. The heatmap's clear organization of variables by their correlation strength aids in identifying the most impactful factors for further analysis. Overall, the results confirm that the dataset's structure supports a robust examination of the factors influencing obesity.
