# Data

## Sources

The dataset utilized in this project was obtained from the UCI Machine Learning Repository, a reputable and extensively used platform for data science and machine learning projects. Originally compiled by researchers at the Universidad de la Costa, Colombia, the dataset combines 77% synthetically generated data with 23% real-world data collected through a structured online survey. The synthetic data, created using the Synthetic Minority Over-sampling Technique (SMOTE) in Weka, addresses class imbalance, enhancing the dataset's suitability for machine learning tasks. The real-world data, gathered from 498 participants over a 30-day period, captures detailed self-reported information on dietary habits, physical activity levels, and demographic characteristics. While synthetic data introduces uniformity and balance, it inherently lacks the complexity of real-world variability, and the user-collected data, though authentic, is susceptible to self-reporting biases and sampling limitations. These characteristics, along with the dataset’s diverse origins, make it an invaluable resource for simulating real-world challenges in healthcare analytics.

## Description

The dataset consists of 2111 records and 17 attributes, offering a detailed examination of the factors contributing to obesity. The attributes represent a mix of categorical and continuous variables, providing insights into demographic, lifestyle, and behavioral factors. In greater detail, an interactive table was designed to provide a comprehensive summary of the dataset's variables.

```{r}
library(here)
library(knitr)
# Main features of the dataset
dataset_raw <- read.csv(here("data/raw/dataset_raw.csv"))
head(dataset_raw) %>%
  kbl(format = "html", caption = "First 6 Rows of the Initial Dataset") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#f0f0f0") %>%
  scroll_box(width = "100%", height = "400px")

desc_table <- tibble(
  Name = colnames(dataset_raw),
  Type = sapply(dataset_raw, class),
  Meaning = val_meaning
)
desc_table %>%
  kbl(format = "html", caption = "Variable Descriptions") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#f0f0f0") %>%
  column_spec(1, bold = TRUE) %>%
  scroll_box(width = "100%", height = "500px")
```

The dataset underwent a thorough preprocessing phase, including normalization of continuous variables, encoding of categorical data, and removal of missing or atypical entries to ensure high-quality analysis. Class imbalance was addressed using the SMOTE (Synthetic Minority Oversampling Technique), generating synthetic data while carefully avoiding noise or artificial patterns. The final dataset comprises 77% synthetic data, which enhances balance and diversity, and 23% real-world data, adding authenticity. This combination allows for a comprehensive analysis of obesity-related factors, while recognizing potential biases, such as inaccuracies in self-reported information.

## Wrangling

To begin the wrangling process, we load the essential libraries required for data manipulation, visualization, and clustering. Each package is selected for its specific functionality, enabling efficient and streamlined analysis:

-   dplyr: for manipulation;

-   tidyr: for data tidying (e.g., reshaping);

-   ggplot2: for visualization;

-   corrplot: for correlation matrix visualization;

-   ggridges: for creating ridge plots;

-   cluster: for clustering algorithms;

-   reshape2: for data reshaping, especially during visualization.

```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(corrplot)
library(ggridges)
library(cluster)
library(reshape2)
```

We rename columns for clarity and ease of use in the analysis. The new names are shorter and more intuitive while preserving their original meaning.

```{r}
  dataset <- dataset_raw %>%
  rename(
    family_hist = family_history_with_overweight,
    obesity_lev = NObeyesdad,
    caloric_food = FAVC,
    vegetable_food = FCVC,
    nb_meal_day = NCP,
    food_btw_meals = CAEC,
    ch2o = CH2O,
    smoke = SMOKE,
    calorie_check = SCC,
    physical_act = FAF,
    freq_alcohol = CALC,
    use_tech = TUE,
    m_trans = MTRANS,
    gender = Gender,
    age = Age,
    weight = Weight,
    height = Height
  )
```

Check for missing values in the dataset, missing values are identified by counting NA values for each column.

```{r}
missing_values <- colSums(is.na(dataset))
missing_values %>%
  kbl(format = "html", caption = "Missing Values in Each Column") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE, position = "center") %>%
  column_spec(1, width = "100px") %>%
  column_spec(2, width = "80px") %>%
  row_spec(0, bold = TRUE, background = "#f0f0f0") %>%
  scroll_box(width = "100%", height = "400px")
```

Missing values are identified by counting NA values for each column. All columns contain complete data, with no missing values. If missing data were present, we could address it by either removing rows with missing values using dataset \<- na.omit(dataset_row) or imputing missing values with appropriate measures (e.g. mean or median).

Check the structure of the dataset to identify data types for each variable. This helps in identifying columns that need to be converted or standardized.

```{r}
# Capture the structure of the dataset
str_output <- capture.output(str(dataset))
# Convert the structure output to a data frame
str_table <- data.frame(Structure = str_output, stringsAsFactors = FALSE)
str_table %>%
  kbl(format = "html", caption = "Structure of the Dataset") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#f0f0f0") %>%
  scroll_box(width = "100%", height = "400px")
```

We convert specific columns to factors for categorical interpretation during analysis. Factors ensure proper handling of discrete variables in statistical modeling.

We arranged the levels of the obesity categories, food consumption between meals, and the frequency of alcohol use to follow a logical ordinal progression, ensuring these variables accurately reflect increasing severity or frequency for improved interpretability and analysis.

```{r}
dataset <- dataset %>%
  mutate(
    gender = as.factor(gender),
    family_hist = as.factor(family_hist),
    caloric_food = as.factor(caloric_food),
    smoke = as.factor(smoke),
    calorie_check = as.factor(calorie_check),
    m_trans = as.factor(m_trans),
    obesity_lev = factor(obesity_lev, 
                         levels = c("Insufficient_Weight", "Normal_Weight", 
                                    "Overweight_Level_I", "Overweight_Level_II", 
                                    "Obesity_Type_I", "Obesity_Type_II", "Obesity_Type_III"), 
                         ordered = TRUE),
    food_btw_meals = factor(ifelse(food_btw_meals == "no", "No", food_btw_meals), 
                            levels = c("No", "Sometimes", "Frequently", "Always"), 
                            ordered = TRUE),
    freq_alcohol = factor(ifelse(freq_alcohol == "no", "No", freq_alcohol), 
                          levels = c("No", "Sometimes", "Frequently", "Always"), 
                          ordered = TRUE))
```

Using str() before and after confirms that each variable has the correct data type, preventing errors during analysis.

```{r}
str_output <- capture.output(str(dataset))
str_table <- data.frame(Structure = str_output, stringsAsFactors = FALSE)
str_table %>%
  kbl(format = "html", caption = "Dataset Structure") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#f5f5f5") %>%
  scroll_box(width = "100%", height = "400px")

```

Check for duplicated rows in the dataset.

```{r}
duplicated_rows <- sum(duplicated(dataset))
duplicated_rows
```

Keep only one instance of each duplicated row.

```{r}
dataset <- dataset %>%
distinct()
```

Check the number of rows after removing duplicates.

```{r}
nrow(dataset)
any(duplicated(dataset))
```

In-depth analysis of SMOTE's impact and visualization of class Distribution

```{r}
library(ggplot2)
g1 <- ggplot(dataset, aes(x = obesity_lev)) +
  geom_bar(fill = "skyblue", color = "black") +
  theme_minimal() +
  labs(
    title = "Class Distribution of Obesity Levels",
    x = "Obesity Level",
    y = "Count"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) #Adjusted the text for clarity

plotly_plot <- ggplotly(g1)
plotly_plot
```

After applying SMOTE, the distribution is noticeably more balanced across all categories, with each class showing a similar count. This outcome reflects SMOTE's intended effect of addressing class imbalance.

Distribution analysis

Density plot for age.

```{r}
g2 <- ggplot(dataset, aes(x = age, fill = obesity_lev)) +
  geom_density(alpha = 0.5) +
  theme_minimal() +
    labs(
    title = "Age Distribution by Obesity Levels",
    x = "Age",
    y = "Density",
    fill = "Obesity Level")
plotly_plot1 <- ggplotly(g2)
plotly_plot1
```

This graph allows us to assess the age distribution across obesity levels and to evaluate the impact of the SMOTE algorithm in generating synthetic data. Two key takeaways emerge: first, the distributions show a clear separation between obesity categories, particularly with younger ages dominating in lower obesity levels (e.g., Insufficient Weight and Normal Weight) and older ages appearing more prominently in higher obesity levels (e.g., Obesity Type II and III). Second, sharp peaks, such as the one around age 30 in "Obesity Type I," could signal potential artifacts from data synthesis. While these patterns indicate that the dataset maintains logical trends, further validation is necessary to confirm that these separations and peaks reflect realistic population characteristics and not artificial biases introduced during data augmentation. Overall, the dataset appears well-structured, but these observations warrant careful consideration during analysis.

Summary statistics by obesity level.

```{r}
dataset_stat <- dataset %>%
  group_by(obesity_lev) %>%
  summarize(
    Age_Mean = mean(age, na.rm = TRUE),
    Age_SD = sd(age, na.rm = TRUE),
    Height_Mean = mean(height, na.rm = TRUE),
    Height_SD = sd(height, na.rm = TRUE),
    Weight_Mean = mean(weight, na.rm = TRUE),
    Weight_SD = sd(weight, na.rm = TRUE)
  )
dataset_stat %>%
  kbl(format = "html", caption = "Summary Statistics by Obesity Level", digits = 1) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#f5f5f5") %>%
  scroll_box(width = "100%", height = "400px")
```

The summary statistics show relatively consistent means and standard deviations for Age, Height, and Weight across obesity levels, which suggests that SMOTE has preserved the overall distribution without introducing extreme values. Interpretation: Since the means and standard deviations are similar across classes, it appears SMOTE didn’t drastically alter the dataset's variability. This consistency supports the idea that SMOTE effectively balanced the classes without distorting key variable distributions.

Perform K-means clustering and calculate silhouette score.

```{r}
library(cluster)
set.seed(123)
kmeans_res <- kmeans(select(dataset, where(is.numeric)), centers = length(unique(dataset$obesity_lev)))
silhouette_score <- silhouette(kmeans_res$cluster, dist(select(dataset, where(is.numeric))))
mean_silhouette_score <- mean(silhouette_score[, "sil_width"])
mean_silhouette_score
```

Silhouette Score from K-means Clustering: The mean silhouette score of approximately 0.456 suggests a moderate level of cohesion within clusters and some separation between them. This score indicates that the clusters (representing obesity levels) are neither too distinct nor too blended. Interpretation: A score close to 0.5 generally reflects reasonable class separability without excessive artificial separability. This score suggests that SMOTE has helped create distinguishable but not overly isolated clusters, which is desirable for class balance. We conclude that SMOTE has balanced the dataset without drastically distorting it.

Creating a Numerical Dataset "dataset_num".

```{r, message=FAlSE, warning=FALSE}

dataset_num <- dataset %>%
  mutate(obesity_lev = recode(obesity_lev,
                              "Insufficient_Weight"=1,
                              "Normal_Weight" = 2,
                              "Overweight_Level_I" = 3,
                              "Overweight_Level_II" = 4,
                              "Obesity_Type_I" = 5,
                              "Obesity_Type_II" = 6,
                              "Obesity_Type_III" = 7,
  ))

dataset_num <- dataset %>%
  mutate(freq_alcohol = recode(freq_alcohol,
                               "No"=1,        
                               "Sometimes"=2, 
                               "Frequently" =3,
                               "Always"  =4 
  ))

dataset_num <- dataset %>%
  mutate(m_trans = recode(m_trans,
                          "Automobile"=1,
                          "Bike"=2,
                          "Motorbike"=3,
                          "Public_Transportation"=4,
                          "Walking"=5,
  ))

dataset_num <- dataset %>%
  mutate(food_btw_meals = recode(food_btw_meals,
                                 "No"=0,
                                 "Sometimes"=1 ,
                                 "Frequently"=2,
                                 "Always"=3
  )
  )

dataset_num <- dataset_num%>%
  mutate(calorie_check = recode(calorie_check,
                                "no"=0,
                                "yes"=1 ,
  ))

dataset_num <- dataset_num %>%
  mutate(across(where(is.factor), ~ as.numeric(.)))


str_output <- capture.output(str(dataset_num))
table_num_str <- data.frame(Structure = str_output, stringsAsFactors = FALSE)

table_num_str %>%
  kbl(format = "html", caption = "Structure of the Numerical Dataset") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#f5f5f5") %>%
  scroll_box(width = "100%", height = "400px")
```

### Spotting Mistakes and Missing Data

We verified the presence of any potential NA values that might have arisen during the conversion of categorical variables to numeric format.

```{r}
nb_na<- colSums(is.na(dataset_num))
nb_na %>%
  kbl(format = "html", caption = "Presence of Potential NA Values in the Dataset") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"), 
    full_width = FALSE, 
    position = "left"
  ) %>%
  column_spec(1, width = "100px") %>%
  column_spec(2, width = "80px") %>%
  row_spec(0, bold = TRUE, background = "#f5f5f5") %>%
  scroll_box(width = "100%", height = "400px")
```

The results of the test confirmed that there are no NA values in the dataset, indicating that all variables were successfully converted to numeric format while retaining their integrity.

### Listing Anomalies and Outliers

## Correlation Analysis

In order to select the possible factor influencing obesity level.

We computed a correlation matrix to analyze the relationships between numeric variables, focusing on their associations with obesity_lev. Variables were reordered by the strength of their correlation with obesity_lev for clarity. A heatmap was generated using a diverging color gradient to visualize these correlations, with red indicating strong positive relationships, blue for negative, and white for weak or neutral. Numerical labels and rotated axis labels were added to improve interpretability, highlighting key factors linked to obesity levels.

```{r}
#Assuming dataset_num is already defined and contains the relevant columns
cor_matrix <- cor(dataset_num %>%
                    select("physical_act", "freq_alcohol", "obesity_lev", "age",
                           "weight","height", "family_hist", "caloric_food",
                           "vegetable_food", "food_btw_meals", "use_tech", "ch2o",
                           "m_trans", "smoke","nb_meal_day", "calorie_check",
                           "gender"),
                  use = "complete.obs")

#Extract the correlations with 'obesity_lev'
cor_with_obesity_lev <- cor_matrix["obesity_lev",]

#Order variables by their correlation with 'obesity_lev'
ordered_vars <- names(sort(cor_with_obesity_lev, decreasing = TRUE))

#Reorder the correlation matrix based on this order
cor_matrix_ordered <- cor_matrix[ordered_vars, ordered_vars]

#Melt the ordered correlation matrix into long format
cor_long <- melt(cor_matrix_ordered)

g3 <- ggplot(cor_long, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() +
  geom_text(aes(label = round(value, 2)), color = "black", size = 2.5, vjust = 0.5
            , hjust = 0.5) + # Center text within tiles
    scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  
  labs(title = "Correlation Heatmap Ordered by Obesity Level", x = "Variables", y
       = "Variables") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        # Rotate x-axis labels for readability
        axis.text.y = element_text(angle = 45, vjust = 1) 
        # Rotate y-axis labels for readability
  )
plot3 <- ggplotly(g3)
plot3
# Create the heatmap with correlation values

# Assuming dataset_num is already defined and contains the relevant columns
cor_matrix <- cor(dataset_num %>%
                    select("physical_act", "freq_alcohol", "obesity_lev", "age",
                           "weight", "family_hist", "caloric_food",
                           "vegetable_food", "food_btw_meals",
                           "use_tech","ch2o", "height",
                           "calorie_check", "gender"),
                  use = "complete.obs")

# Extract the correlations with 'obesity_lev'
cor_with_obesity_lev <- cor_matrix["obesity_lev",]

# Order variables by their correlation with 'obesity_lev'
ordered_vars <- names(sort(cor_with_obesity_lev, decreasing = TRUE))

# Reorder the correlation matrix based on this order
cor_matrix_ordered <- cor_matrix[ordered_vars, ordered_vars]

# Melt the ordered correlation matrix into long format
cor_long <- melt(cor_matrix_ordered)


g4 <- ggplot(cor_long, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() +
  geom_text(aes(label = round(value, 2)), color = "black", size = 2.5, vjust = 0.5
            , hjust = 0.5) + # Center text within tiles
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  labs(title = "Correlation Heatmap Ordered by Obesity Level", x = "Variables", y
       = "Variables") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        # Rotate x-axis labels for readability
        axis.text.y = element_text(angle = 45, vjust = 1) 
        # Rotate y-axis labels for readability
  )
plot4 <- ggplotly(g4)
plot4
```
