# Data

## Sources

The dataset utilized in this project was obtained from the UCI Machine Learning Repository, a reputable and extensively used platform for data science and machine learning projects. Originally compiled by researchers at the Universidad de la Costa, Colombia, the dataset combines 77% synthetically generated data with 23% real-world data collected through a structured online survey. The synthetic data, created using the Synthetic Minority Over-sampling Technique (SMOTE) in Weka, addresses class imbalance, enhancing the dataset's suitability for machine learning tasks. The real-world data, gathered from 498 participants over a 30-day period, captures detailed self-reported information on dietary habits, physical activity levels, and demographic characteristics. While synthetic data introduces uniformity and balance, it inherently lacks the complexity of real-world variability, and the user-collected data, though authentic, is susceptible to self-reporting biases and sampling limitations. These characteristics, along with the dataset’s diverse origins, make it an invaluable resource for simulating real-world challenges in healthcare analytics.

## Description

The dataset consists of 2111 records and 17 attributes, offering a detailed examination of the factors contributing to obesity. The attributes represent a mix of categorical and continuous variables, providing insights into demographic, lifestyle, and behavioral factors. In greater detail, an interactive table was designed to provide a comprehensive summary of the dataset's variables.

```{r}
library(here)
library(knitr)
# Main features of the dataset
dataset_raw <- read.csv(here("data/raw/dataset_raw.csv"))
head(dataset_raw) %>%
  kbl(format = "html", caption = "First 6 Rows of the Initial Dataset") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#f0f0f0") %>%
  scroll_box(width = "100%", height = "400px")

desc_table <- tibble(
  Name = colnames(dataset_raw),
  Type = sapply(dataset_raw, class),
  Meaning = val_meaning
)
desc_table %>%
  kbl(format = "html", caption = "Variable Descriptions") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#f0f0f0") %>%
  column_spec(1, bold = TRUE) %>%
  scroll_box(width = "100%", height = "500px")
```

The dataset underwent a thorough preprocessing phase, including normalization of continuous variables, encoding of categorical data, and removal of missing or atypical entries to ensure high-quality analysis. Class imbalance was addressed using the SMOTE (Synthetic Minority Oversampling Technique), generating synthetic data while carefully avoiding noise or artificial patterns. The final dataset comprises 77% synthetic data, which enhances balance and diversity, and 23% real-world data, adding authenticity. This combination allows for a comprehensive analysis of obesity-related factors, while recognizing potential biases, such as inaccuracies in self-reported information.

## Wrangling

To initiate the wrangling process, essential libraries for data manipulation, visualization, and clustering are loaded. Each package is utilized for its specific functionality, facilitating efficient and streamlined analysis:

-   dplyr: for data manipulation (e.g., filtering, summarizing);

-   tidyr: for data tidying (e.g., reshaping);

-   ggplot2: for visualization;

-   corrplot: for correlation matrix visualization;

-   ggridges: for creating ridge plots;

-   cluster: for clustering algorithms;

-   reshape2: for data reshaping, especially during visualization.

```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(corrplot)
library(ggridges)
library(cluster)
library(reshape2)
```

Column names are renamed to enhance clarity and improve usability during the analysis. The updated names are designed to be shorter and more intuitive, ensuring ease of reference while retaining their original meaning and context. This adjustment simplifies code readability and helps streamline data manipulation tasks, particularly in complex analytical workflows.

```{r}
  dataset <- dataset_raw %>%
  rename(
    family_hist = family_history_with_overweight,
    obesity_lev = NObeyesdad,
    caloric_food = FAVC,
    vegetable_food = FCVC,
    nb_meal_day = NCP,
    food_btw_meals = CAEC,
    ch2o = CH2O,
    smoke = SMOKE,
    calorie_check = SCC,
    physical_act = FAF,
    freq_alcohol = CALC,
    use_tech = TUE,
    m_trans = MTRANS,
    gender = Gender,
    age = Age,
    weight = Weight,
    height = Height
  )
```

**Check for missing values**

To ensure data integrity, missing values in the dataset are examined by counting "NA" values in each column, providing a clear view of dataset completeness. The results are presented in a formatted table for easy interpretation, with styling applied for readability and a scrollable box to handle larger datasets. This process facilitates prompt handling of missing data through appropriate strategies.

```{r}
missing_values <- colSums(is.na(dataset))
missing_values %>%
  kbl(format = "html", caption = "Missing Values in Each Column") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE, position = "center") %>%
  column_spec(1, width = "100px") %>%
  column_spec(2, width = "80px") %>%
  row_spec(0, bold = TRUE, background = "#f0f0f0") %>%
  scroll_box(width = "100%", height = "400px")
```

The analysis confirms that all columns contain complete data, with no missing values identified. This completeness ensures a robust foundation for subsequent analysis, eliminating the need for immediate data cleaning related to missing entries.

**Check for dataset structure**

The structure of the dataset is examined to identify the data types of each variable, providing critical insights for subsequent data preparation. Understanding the data types helps pinpoint columns requiring transformations, such as converting categorical variables to factors or standardizing numeric variables for analysis.

```{r}
str_output <- capture.output(str(dataset))
str_table <- data.frame(Structure = str_output, stringsAsFactors = FALSE)
str_table %>%
  kbl(format = "html", caption = "Original structure of the Dataset") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#f0f0f0") %>%
  scroll_box(width = "100%", height = "400px")


dataset <- dataset %>%
  mutate(
    gender = as.factor(gender),
    family_hist = as.factor(family_hist),
    caloric_food = as.factor(caloric_food),
    smoke = as.factor(smoke),
    calorie_check = as.factor(calorie_check),
    m_trans = as.factor(m_trans),
    obesity_lev = factor(obesity_lev, 
                         levels = c("Insufficient_Weight", "Normal_Weight", 
                                    "Overweight_Level_I", "Overweight_Level_II", 
                                    "Obesity_Type_I", "Obesity_Type_II", "Obesity_Type_III"), 
                         ordered = TRUE),
    food_btw_meals = factor(ifelse(food_btw_meals == "no", "No", food_btw_meals), 
                            levels = c("No", "Sometimes", "Frequently", "Always"), 
                            ordered = TRUE),
    freq_alcohol = factor(ifelse(freq_alcohol == "no", "No", freq_alcohol), 
                          levels = c("No", "Sometimes", "Frequently", "Always"), 
                          ordered = TRUE))


str_output <- capture.output(str(dataset))
str_table <- data.frame(Structure = str_output, stringsAsFactors = FALSE)
str_table %>%
  kbl(format = "html", caption = "Dataset Structure") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#f5f5f5") %>%
  scroll_box(width = "100%", height = "400px")
```

The transformations ensured the dataset was ready for analysis by restructuring categorical and ordinal variables to meet modeling requirements. Converting categorical variables into factors standardized their representation, reducing ambiguity and improving compatibility with statistical models. For ordinal variables, levels were explicitly ordered to preserve their logical progression and enhance interpretability, allowing for meaningful comparisons across categories.

The updated structure was reviewed to confirm the accuracy of these adjustments, providing confidence in the preprocessing steps. While further transformations like normalization were not applied, the focus on categorical and ordinal adjustments established a strong foundation for reliable and interpretable analysis. In particular, the levels of obesity categories, food consumption between meals, and frequency of alcohol use were arranged to reflect increasing severity or frequency, ensuring these variables captured their intended relationships and supported clear, accurate insights into the data.

**Check for duplicates**

The dataset is examined for duplicated rows to ensure data integrity and eliminate redundancy. Identifying and addressing duplicates is a crucial step in data preprocessing, as redundant entries can skew analysis results and lead to misleading conclusions. This process involves systematically scanning the dataset for identical rows and quantifying their occurrence.

```{r}
duplicated_rows <- sum(duplicated(dataset))
duplicated_rows
```

Keep only one instance of each duplicated row.

```{r}
dataset <- dataset %>%
distinct()

nrow(dataset)
any(duplicated(dataset))
```

Check the number of rows after removing duplicates.

**SMOTE's analysis**

In-depth analysis of SMOTE's impact and visualization of class Distribution

```{r}
library(ggplot2)
g1 <- ggplot(dataset, aes(x = obesity_lev)) +
  geom_bar(fill = "skyblue", color = "black") +
  theme_minimal() +
  labs(
    title = "Class Distribution of Obesity Levels",
    x = "Obesity Level",
    y = "Count"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) #Adjusted the text for clarity

plotly_plot <- ggplotly(g1)
plotly_plot
```

After applying SMOTE, the distribution is noticeably more balanced across all categories, with each class showing a similar count. This outcome reflects SMOTE's intended effect of addressing class imbalance.

Distribution analysis

Density plot for age.

```{r}
g2 <- ggplot(dataset, aes(x = age, fill = obesity_lev)) +
  geom_density(alpha = 0.5) +
  theme_minimal() +
    labs(
    title = "Age Distribution by Obesity Levels",
    x = "Age",
    y = "Density",
    fill = "Obesity Level")
plotly_plot1 <- ggplotly(g2)
plotly_plot1
```

This graph allows us to assess the age distribution across obesity levels and to evaluate the impact of the SMOTE algorithm in generating synthetic data. Two key takeaways emerge: first, the distributions show a clear separation between obesity categories, particularly with younger ages dominating in lower obesity levels (e.g., Insufficient Weight and Normal Weight) and older ages appearing more prominently in higher obesity levels (e.g., Obesity Type II and III). Second, sharp peaks, such as the one around age 30 in "Obesity Type I," could signal potential artifacts from data synthesis. While these patterns indicate that the dataset maintains logical trends, further validation is necessary to confirm that these separations and peaks reflect realistic population characteristics and not artificial biases introduced during data augmentation. Overall, the dataset appears well-structured, but these observations warrant careful consideration during analysis.

Summary statistics by obesity level.

```{r}
dataset_stat <- dataset %>%
  group_by(obesity_lev) %>%
  summarize(
    Age_Mean = mean(age, na.rm = TRUE),
    Age_SD = sd(age, na.rm = TRUE),
    Height_Mean = mean(height, na.rm = TRUE),
    Height_SD = sd(height, na.rm = TRUE),
    Weight_Mean = mean(weight, na.rm = TRUE),
    Weight_SD = sd(weight, na.rm = TRUE)
  )
dataset_stat %>%
  kbl(format = "html", caption = "Summary Statistics by Obesity Level", digits = 1) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#f5f5f5") %>%
  scroll_box(width = "100%", height = "400px")
```

The summary statistics show relatively consistent means and standard deviations for Age, Height, and Weight across obesity levels, which suggests that SMOTE has preserved the overall distribution without introducing extreme values. Interpretation: Since the means and standard deviations are similar across classes, it appears SMOTE didn’t drastically alter the dataset's variability. This consistency supports the idea that SMOTE effectively balanced the classes without distorting key variable distributions.

Perform K-means clustering and calculate silhouette score.

```{r}
library(cluster)
set.seed(123)
kmeans_res <- kmeans(select(dataset, where(is.numeric)), centers = length(unique(dataset$obesity_lev)))
silhouette_score <- silhouette(kmeans_res$cluster, dist(select(dataset, where(is.numeric))))
mean_silhouette_score <- mean(silhouette_score[, "sil_width"])
mean_silhouette_score
```

Silhouette Score from K-means Clustering: The mean silhouette score of approximately 0.456 suggests a moderate level of cohesion within clusters and some separation between them. This score indicates that the clusters (representing obesity levels) are neither too distinct nor too blended. Interpretation: A score close to 0.5 generally reflects reasonable class separability without excessive artificial separability. This score suggests that SMOTE has helped create distinguishable but not overly isolated clusters, which is desirable for class balance. We conclude that SMOTE has balanced the dataset without drastically distorting it.

Creating a Numerical Dataset "dataset_num".

```{r, message=FAlSE, warning=FALSE}

dataset_num <- dataset %>%
  mutate(obesity_lev = recode(obesity_lev,
                              "Insufficient_Weight"=1,
                              "Normal_Weight" = 2,
                              "Overweight_Level_I" = 3,
                              "Overweight_Level_II" = 4,
                              "Obesity_Type_I" = 5,
                              "Obesity_Type_II" = 6,
                              "Obesity_Type_III" = 7,
  ))

dataset_num <- dataset %>%
  mutate(freq_alcohol = recode(freq_alcohol,
                               "No"=1,        
                               "Sometimes"=2, 
                               "Frequently" =3,
                               "Always"  =4 
  ))

dataset_num <- dataset %>%
  mutate(m_trans = recode(m_trans,
                          "Automobile"=1,
                          "Bike"=2,
                          "Motorbike"=3,
                          "Public_Transportation"=4,
                          "Walking"=5,
  ))

dataset_num <- dataset %>%
  mutate(food_btw_meals = recode(food_btw_meals,
                                 "No"=0,
                                 "Sometimes"=1 ,
                                 "Frequently"=2,
                                 "Always"=3
  )
  )

dataset_num <- dataset_num%>%
  mutate(calorie_check = recode(calorie_check,
                                "no"=0,
                                "yes"=1 ,
  ))

dataset_num <- dataset_num %>%
  mutate(across(where(is.factor), ~ as.numeric(.)))


str_output <- capture.output(str(dataset_num))
table_num_str <- data.frame(Structure = str_output, stringsAsFactors = FALSE)

table_num_str %>%
  kbl(format = "html", caption = "Structure of the Numerical Dataset") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#f5f5f5") %>%
  scroll_box(width = "100%", height = "400px")
```

### Spotting Mistakes and Missing Data

We verified the presence of any potential NA values that might have arisen during the conversion of categorical variables to numeric format.

```{r}
nb_na<- colSums(is.na(dataset_num))
nb_na %>%
  kbl(format = "html", caption = "Presence of Potential NA Values in the Dataset") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"), 
    full_width = FALSE, 
    position = "left"
  ) %>%
  column_spec(1, width = "100px") %>%
  column_spec(2, width = "80px") %>%
  row_spec(0, bold = TRUE, background = "#f5f5f5") %>%
  scroll_box(width = "100%", height = "400px")
```

The results of the test confirmed that there are no NA values in the dataset, indicating that all variables were successfully converted to numeric format while retaining their integrity.

### Listing Anomalies and Outliers

## Correlation Analysis

In order to select the possible factor influencing obesity level.

We computed a correlation matrix to analyze the relationships between numeric variables, focusing on their associations with obesity_lev. Variables were reordered by the strength of their correlation with obesity_lev for clarity. A heatmap was generated using a diverging color gradient to visualize these correlations, with red indicating strong positive relationships, blue for negative, and white for weak or neutral. Numerical labels and rotated axis labels were added to improve interpretability, highlighting key factors linked to obesity levels.

```{r}
#Assuming dataset_num is already defined and contains the relevant columns
cor_matrix <- cor(dataset_num %>%
                    select("physical_act", "freq_alcohol", "obesity_lev", "age",
                           "weight","height", "family_hist", "caloric_food",
                           "vegetable_food", "food_btw_meals", "use_tech", "ch2o",
                           "m_trans", "smoke","nb_meal_day", "calorie_check",
                           "gender"),
                  use = "complete.obs")

#Extract the correlations with 'obesity_lev'
cor_with_obesity_lev <- cor_matrix["obesity_lev",]

#Order variables by their correlation with 'obesity_lev'
ordered_vars <- names(sort(cor_with_obesity_lev, decreasing = TRUE))

#Reorder the correlation matrix based on this order
cor_matrix_ordered <- cor_matrix[ordered_vars, ordered_vars]

#Melt the ordered correlation matrix into long format
cor_long <- melt(cor_matrix_ordered)

g3 <- ggplot(cor_long, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() +
  geom_text(aes(label = round(value, 2)), color = "black", size = 2.5, vjust = 0.5
            , hjust = 0.5) + # Center text within tiles
    scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  
  labs(title = "Correlation Heatmap Ordered by Obesity Level", x = "Variables", y
       = "Variables") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        # Rotate x-axis labels for readability
        axis.text.y = element_text(angle = 45, vjust = 1) 
        # Rotate y-axis labels for readability
  )
plot3 <- ggplotly(g3)
plot3
# Create the heatmap with correlation values

# Assuming dataset_num is already defined and contains the relevant columns
cor_matrix <- cor(dataset_num %>%
                    select("physical_act", "freq_alcohol", "obesity_lev", "age",
                           "weight", "family_hist", "caloric_food",
                           "vegetable_food", "food_btw_meals",
                           "use_tech","ch2o", "height",
                           "calorie_check", "gender"),
                  use = "complete.obs")

# Extract the correlations with 'obesity_lev'
cor_with_obesity_lev <- cor_matrix["obesity_lev",]

# Order variables by their correlation with 'obesity_lev'
ordered_vars <- names(sort(cor_with_obesity_lev, decreasing = TRUE))

# Reorder the correlation matrix based on this order
cor_matrix_ordered <- cor_matrix[ordered_vars, ordered_vars]

# Melt the ordered correlation matrix into long format
cor_long <- melt(cor_matrix_ordered)


g4 <- ggplot(cor_long, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() +
  geom_text(aes(label = round(value, 2)), color = "black", size = 2.5, vjust = 0.5
            , hjust = 0.5) + # Center text within tiles
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  labs(title = "Correlation Heatmap Ordered by Obesity Level", x = "Variables", y
       = "Variables") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        # Rotate x-axis labels for readability
        axis.text.y = element_text(angle = 45, vjust = 1) 
        # Rotate y-axis labels for readability
  )
plot4 <- ggplotly(g4)
plot4
```
