# 3. Exploratory Data Analysis (EDA)

#### 3.1 Descriptive statistics and distribution analysis

##### Age

Descriptive statistic for Age

```{r}
# this is what was written before in case you don't like the new output 
#summary(dataset$age)
# sd(dataset$age, na.rm = TRUE)
sum_age_df <- tibble::tibble(
  Metric = c(names(summary(dataset$age)), "Std. Dev"),
  Value = c(summary(dataset$age), sd(dataset$age, na.rm = TRUE))
)
kable(sum_age_df, format = "markdown", caption = "Summary of the age variable")

```

Age distribution

The age data shows a right-skewed distribution, with a mean of 24.3 years and a median of 22.78 years. The range (14 to 61 years) covers a wide age span, but most individuals are concentrated in the 20–30 age range. The standard deviation (6.35 years) suggests moderate variability in the dataset. This young population distribution may limit the applicability of results to older age groups, where obesity risk factors could differ.

Age Distribution by Obesity Level (Violin Plot)

The age distribution varies across obesity levels,highlighting distinct trends. Insufficient and normal weight categories are concentrated among younger individuals (14–30), while overweight and obesity levels shift towards mid-adulthood (20–40), peaking around 30–35 years. Severe obesity (Type III) is rare in younger ages and more common in the 30–40 range. These patterns suggest the progression of weight issues with age and emphasize the need for targeted interventions during early to mid-adulthood to prevent worsening obesity levels.

```{r}
ggplot(dataset, aes(x = obesity_lev, y = age, fill = obesity_lev)) +
  geom_violin(trim = FALSE, alpha = 0.6) +
  geom_boxplot(width = 0.1, color = "black", fill = "white") +
  labs(title = "Age Distribution by Obesity Level", x = "Obesity Level", y = "Age") +
  theme_minimal() +
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

The violin plot shows, more clearly, how individuals in the lower obesity categories, such as insufficient and normal weight, are predominantly younger, with ages concentrated between 14 and 30 years. In contrast, higher obesity levels exhibit a broader age range, with a peak density observed around 30–40 years, particularly in Obesity Type I and Type II. Severe obesity (Type III) is rare in younger individuals and becomes more prominent in the mid-adulthood age group. This visualization underscores the gradual progression of obesity risk with age and emphasizes the critical need for early intervention strategies to address weight-related health issues, particularly during early and mid-adulthood when such risks become more pronounced.

Age Distribution with SMOOTH Trend Line for Obesity Probability.

```{r, message=FALSE, warning=FALSE}
ggplot(dataset, aes(x = age, y = as.numeric(obesity_lev))) +
  geom_jitter(alpha = 0.3) +
  geom_smooth(method = "loess", se = FALSE, color = "blue") +
  labs(title = "Trend of Obesity Level with Age", x = "Age", y = "Obesity Level") +
  theme_minimal()
```

The graph shows a smooth trend line capturing the overall pattern. Obesity levels increase significantly from adolescence to early adulthood, peaking around the 25–30 years age range. This period potentially represents a critical transition, where lifestyle factors such as reduced physical activity, higher caloric intake, and metabolic changes can contribute to the steep rise in obesity levels.

Beyond the peak, the trend shows a gradual decline in obesity levels after 30 years, which may reflect behavioral changes, such as increased health awareness, dietary improvements, or a selection bias in older age groups. This switch suggests that mid-20s to early-30s is a pivotal stage for interventions aimed at mitigating obesity risk.

##### Height

Descriptive statistic for Height.

```{r}
# summary(dataset$height)
# sd(dataset$height, na.rm = TRUE)

sum_height_df <- tibble::tibble(
  Metric = c(names(summary(dataset$height)), "Std. Dev"),
  Value = c(summary(dataset$height), sd(dataset$height, na.rm = TRUE))
)
kable(sum_height_df, format = "markdown", caption = "Summary of the height variable")
```

Height distribution.

```{r}
ggplot(dataset, aes(x = height)) +
  geom_histogram(bins = 20, fill = "purple", color = "black", alpha = 0.7) +
  labs(title = "Height Distribution", x = "Height (m)", y = "Count") +
  theme_minimal()
```

The height histogram shows the height distribution (in meters) and is approximately normal, with a slight right skew. Most values fall between 1.45m and 1.98m, with a peak around 1.8m, indicating it’s the most frequent height. The range is realistic, with no visible extreme outliers, and the standard deviation (0.09) indicates low variability. I would like to add that the mean and median are both 1.7m, confirming a nearly symmetrical distribution.

Height by Obesity Level

Box Plot of Height by Obesity Level.

```{r}
ggplot(dataset, aes(x = obesity_lev, y = height, fill = obesity_lev)) +
  geom_violin(alpha = 0.6) +
  labs(title = "Height Distribution by Obesity Level", x = "Obesity Level", y = "Height") +
  theme_minimal() +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))

```

The plot shows for height, relatively low variability within each category, with overlapping ranges between most groups. Individuals with Insufficient Weight and Normal Weight have slightly narrower distributions, centered around similar heights (\~1.7 m). As obesity levels increase (e.g., Obesity Type I–III), the distributions remain consistent, suggesting height is not strongly associated with obesity classification. This suggests that weight may be more influential than height alone in determining obesity level.

##### Weight

Descriptive statistic for Weight.

```{r}
# summary(dataset$weight)
# sd(dataset$weight, na.rm = TRUE)

sum_weight_df <- tibble::tibble(
  Metric = c(names(summary(dataset$weight)), "Std. Dev"),
  Value = c(summary(dataset$weight), sd(dataset$weight, na.rm = TRUE))
)
kable(sum_weight_df, format = "markdown", caption = "Summary of the weight variable")
```

Weight by gender

Density plot for weight distribution by gender.

```{r}
ggplot(dataset, aes(x = weight, fill = gender)) +
  geom_density(alpha = 0.5) +
  labs(title = "Density Plot of Weight by Gender", x = "Weight", y = "Density") +
  scale_fill_manual(values = c("pink", "lightblue"), name = "Gender", labels = c("Female", "Male")) +
  theme_minimal()
```

The density plot reveals distinct weight distributions between genders. Females generally weight less, with a peak around 70 units, while males peak around 85 and 115 units, indicating a tendency toward higher weights. The overlapping region around 80-90 units shows weights common to both genders, but the distinct density peaks emphasize gender-based differences in weight distribution. Overall, males dominate at higher ranges Weight ranges from 39 to 173 units, with an average (mean) weight of 86.6 units. The median weight is 83 units, with a standard deviation of 26.2, indicating moderate spread.

Weight by obesity level

Ridgeline Plot of Weight by Obesity Level.

```{r, message=FALSE, warning=FALSE}
ggplot(dataset, aes(x = weight, y = obesity_lev, fill = obesity_lev)) +
  geom_density_ridges(scale = 0.9, alpha = 0.6) +
  labs(title = "Ridgeline Plot of Weight by Obesity Level", x = "Weight", y = "Obesity Level") +
  theme_minimal() +
  theme(legend.position = "none")
```

This ridgeline plot shows a clear progression in weight distribution across different obesity levels. As the obesity level increases, the weight distribution shifts progressively to higher ranges. "Normal Weight" and "Insufficient Weight" categories are concentrated at lower weights, while higher obesity types (I, II, and III) peak at significantly greater weights, indicating a strong positive association between weight and obesity level The weight distribution has an average of 86.6 kg and a standard deviation of 26.6 kg.

##### Height and Weight

Scatter Plot (height vs weight), colored by obesity level.

```{r, message=FALSE, warning=FALSE}
ggplot(dataset, aes(x = height, y = weight, color = obesity_lev)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, aes(group = obesity_lev)) +  # Adds a trend line for each obesity level
  ggtitle("Scatter Plot of Weight vs Height by Obesity Level") +
  theme_minimal() +
  labs(x = "Height", y = "Weight", color = "Obesity Level")
```

Facet Grid for Height and Weight by Obesity Level.

```{r}
ggplot(dataset, aes(x = height, y = weight)) +
  geom_point(alpha = 0.7, aes(color = obesity_lev)) +
  facet_wrap(~ obesity_lev) +
  ggtitle("Facet Grid of Weight and Height by Obesity Level") +
  theme_minimal() +
  labs(x = "Height", y = "Weight", color = "Obesity Level") +
  theme(legend.position = "none")
```

The scatter plot with trend lines for each obesity level reveals a clear positive correlation between weight and height across all obesity levels. As the obesity level increases, the slope generally becomes steeper, indicating a stronger weight gain relative to height. We created the facet grid to show more clearly the trends to show more clearly how The "Obesity_Type_III" (yellow) category has the steepest slope, suggesting a significant weight increase per unit of height, which is consistent with the highest level of obesity.

Correlation between height and weight.

```{r}
correlation_height_weight <- cor(dataset$height, dataset$weight, use = "complete.obs")
correlation_height_weight
```

The correlation observed between height and weight (r = 0.463) aligns with existing literature, confirming the expected positive relationship between these variables.

##### Food between meals

```{r}

# Dodged Bar Chart for food_btw_meals by obesity levels
ggplot(dataset, aes(x = food_btw_meals, fill = obesity_lev)) +
   geom_bar(position = "dodge", color = "black") +
   ggtitle("Dodged Bar Chart for Food Between Meals by Obesity Levels") +
   labs(x = "Food Between Meals", y = "Count", fill = "Obesity Levels") +
   theme_minimal() +
   theme(
         plot.title = element_text(hjust = 0.5, size = 14))

# Stacked Bar Chart of Food Between Meals by Obesity Level (Proportions within each Obesity Level)
ggplot(dataset, aes(x = obesity_lev, fill = food_btw_meals)) +
    geom_bar(position = "fill") + # Stacked bar chart with proportions
    scale_y_continuous(labels = scales::percent_format(accuracy = 1)) + # Format y-axis as percentages
    ggtitle("Proportion of Food Between Meals Across Obesity Levels") + # Shortened and clear title
    labs(x = "Obesity Levels", y = "Proportion (%)", fill = "Food Between Meals") + # Correct axis and legend labels
    theme_minimal() +
    theme(
        axis.text.x = element_text(angle = 45, hjust = 1), # Rotate x-axis text for readability
        plot.title = element_text(hjust = 0.5, size = 14) # Center and style the title
    )
```

These charts provide a clear illustration of how the frequency of eating between meals varies across obesity levels. The most dominant behavior across all categories is "Sometimes," which peaks in intermediate levels like Normal Weight and Overweight Level I, reflecting a common pattern of moderate snacking. However, as obesity levels increase to Obesity Types I–III, the responses for "Frequently" and "Always" diminish, while "Sometimes" becomes even more prevalent. This shift could indicate that higher obesity levels are more associated with habitual moderate snacking rather than excessive meal-snacking frequency. On the other hand, "No" responses remain negligible across all obesity levels, suggesting that eating between meals is almost universal in this population. This pattern underscores the importance of examining not just the frequency but also the quality and context of snacking as potential contributors to obesity progression.

##### High-caloric food consumption

```{r}
# Dodged Bar Chart for High-Caloric Food Consumption by Obesity Levels
ggplot(dataset, aes(x = caloric_food, fill = obesity_lev)) +
   geom_bar(position = "dodge", color = "black") +
   ggtitle("    Dodged Bar Chart for High-Caloric Food Consumption by Obesity Levels") +
   labs(x = "High-Caloric Food Consumption", y = "Count", fill = "Obesity Levels") +
   theme_minimal() +
   theme(
         plot.title = element_text(hjust = 0.5, size = 14)) # Center and style the title
```

The dodged bar chart clearly shows that the majority of individuals, especially in the higher obesity categories (Obesity Type I–III), report consuming high-caloric foods ("yes"). This trend becomes increasingly pronounced as obesity levels rise, with very few individuals reporting "no" consumption in these categories. In contrast, lower obesity levels (e.g., Normal Weight, Overweight Level I) show a slightly higher representation of "no" responses, indicating a potential shift in dietary habits across obesity levels.

```{r}
# Grouped Bar Chart of High-Caloric Food by Obesity Level (Proportions within each Obesity Level)
ggplot(dataset, aes(x = obesity_lev, fill = caloric_food)) +
  geom_bar(
    position = "dodge",
    aes(y = after_stat(count) / tapply(after_stat(count), after_stat(x), sum)[after_stat(x)]),
    color = "black") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  ggtitle("                                 Grouped Bar Chart of High-Caloric Food Consumption Across Obesity Levels") +
  labs(x = "Obesity Levels", y = "Proportion (%)", fill = "High-Caloric Food Consumption") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5, size = 14)
  )

```

The grouped bar chart effectively shows the behavioral shift toward higher high-caloric food consumption as obesity levels increase. High-caloric food consumption ("yes") consistently accounts for over 75% of responses, becoming nearly universal in higher obesity categories (Obesity Type I–III). In contrast, "no" responses are more visible in lower obesity levels, such as Insufficient Weight and Normal Weight, but remain a minority.

```{r}
percentage_high_caloric_consumers <- mean(dataset$caloric_food == "yes") * 100
percentage_high_caloric_consumers
```

More precisely, a notable 88.4% of participants report frequent consumption of high-calorie foods, which may directly contribute to weight gain, highlighting the need for dietary interventions focused on reducing high-calorie intake.

##### Alcohol consumption

Frequence in consumption of alcohol.

```{r}
# Filter out "Always" responses from the dataset
filtered_dataset <- dataset %>%
  filter(freq_alcohol != "Always")

# Dodged Bar Chart for freq_alcohol by Obesity Levels (excluding "Always")
ggplot(filtered_dataset, aes(x = freq_alcohol, fill = obesity_lev)) +
   geom_bar(position = "dodge", color = "black") +
   ggtitle("Dodged Bar Chart for Alcohol Consumption by Obesity Levels") +
   labs(x = "Alcohol Consumption Frequency", y = "Count", fill = "Obesity Levels") +
   theme_minimal() +
   theme(
         plot.title = element_text(hjust = 0.5, size = 14)) # Center and style the title




```

The chart shows that "Sometimes" is the dominant alcohol consumption frequency across all obesity levels, particularly in Normal Weight, Overweight Level I, and II categories. As obesity increases, "Frequently" becomes slightly more prominent, especially in Obesity Type III, while "No" responses decrease, being more common in lower obesity levels such as Insufficient and Normal Weight. The "Always" responses are excluded from this chart due to their near absence in the dataset, highlighting that excessive alcohol consumption is rare. This trend underlines the potential relationship between moderate-to-frequent alcohol consumption and higher obesity levels, emphasizing its importance for obesity-related behavioral research.

```{r}
# Prepare the data summary for 'Sometimes' and 'No' responses
data_summary <- dataset %>%
  filter(freq_alcohol %in% c("Sometimes", "No")) %>%
  group_by(obesity_lev, freq_alcohol) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(obesity_lev) %>%
  mutate(
    total = sum(count),
    proportion = count / total
  ) %>%
  ungroup()

# Visualization with updated title
ggplot(data_summary, aes(x = obesity_lev, y = proportion, group = freq_alcohol, color = freq_alcohol)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +  # Format y-axis as percentages
  ggtitle("Proportion of 'Sometimes' and 'No' Alcohol Responses by Obesity Level") +
  labs(x = "Obesity Level", y = "Proportion (%)", color = "Alcohol Frequency") +
  scale_color_manual(values = c("No" = "purple", "Sometimes" = "gold")) + # Improved color scheme
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5, size = 14),  # Center and style title
    legend.position = "top"
  )
```

The proportion of individuals who drink alcohol "Sometimes" increases with higher obesity levels, peaking in Obesity_Type_III. In contrast, the likelihood of abstaining from alcohol ("no") decreases as obesity levels rise. This pattern suggests that moderate alcohol consumption may be associated with higher obesity levels, while abstention is more common among those with lower obesity levels.

A possible interaction to investigate later is between alcohol frequency and caloric food preference, as both behaviors seem linked to higher obesity levels. Exploring this could reveal if individuals with a preference for caloric foods and moderate alcohol consumption have a compounding effect on obesity risk. This investigation could help clarify whether combined lifestyle factors contribute more significantly to higher obesity levels than each factor alone.

Monitoring of the calories in the day.

```{r}
# Dodged Bar Chart for calorie_check by Obesity Levels
ggplot(dataset, aes(x = calorie_check, fill = obesity_lev)) +
   geom_bar(position = "dodge", color = "black") +
   ggtitle("    Dodged Bar Chart for the check of the calories by Obesity Levels") +
   labs(x = "High-Caloric Food Consumption", y = "Count", fill = "Obesity Levels") +
   theme_minimal() +
   theme(
         plot.title = element_text(hjust = 0.5, size = 14)) # Center and style the title


data_summary <- dataset %>%
  group_by(obesity_lev, calorie_check) %>%
  summarise(count = n(), .groups = "drop") %>%
  mutate(total = sum(count), proportion = count / total)

ggplot(data_summary, aes(x = obesity_lev, y = proportion, group = calorie_check, color = calorie_check)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  scale_y_continuous(labels = scales::percent) +
  scale_color_manual(values = c("no" = "lightcoral", "yes" = "lightblue")) +
  labs(title = "Proportion of Calorie Checking by Obesity Level", x = "Obesity Level", y = "Proportion", color = "Calorie Check") +
  theme_minimal() +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))

```

The Dodged Bar Chart highlights two main trends regarding calorie-checking behavior across obesity levels: a significant increase in "Yes" responses as obesity levels rise, particularly from Overweight Level II onward, and a decrease in "No" responses, which are more prevalent in lower obesity levels like Normal Weight and Insufficient Weight. The second graph simplifies these trends by clearly illustrating the proportional shift between "Yes" and "No" responses, making the contrast between lower and higher obesity levels more visually apparent. Together, these visualizations emphasize a potential association between obesity severity and an increased tendency to check calorie intake, suggesting heightened dietary awareness in higher obesity categories.

##### Vegetable consumption

```{r}
ggplot(dataset, aes(x = vegetable_food)) +
  geom_histogram(aes(y =after_stat(density)), bins = 30, fill = "lightgreen", color = "black", alpha = 0.6) +
  geom_density(color = "darkgreen", linewidth = 1) +
  ggtitle("Histogram and Density of Vegetable Food Consumption") +
  theme_minimal() +
  labs(x = "Vegetable Food Consumption", y = "Density")
```

```{r, message=FALSE, warning=FALSE}
ggplot(dataset, aes(x = weight, y = vegetable_food, color = obesity_lev)) +
    geom_point(alpha = 0.6) +
    geom_smooth(method = "loess", se = FALSE, color = "black") +
    labs(title = "Scatterplot of Weight vs Vegetable Food Consumption", 
         x = "Weight", 
         y = "Vegetable Food Consumption") +
    theme_minimal() +
    coord_cartesian(xlim= c(40, 135), ylim= c(2, 3))
```

The scatterplot provided with the trend line illustrates a distinct, non-linear relationship: vegetable consumption initially decreases as weight increases but then begins to rise again at higher weight levels.

This pattern suggests that individuals with lower weight, particularly those in the Insufficient Weight and Normal Weight categories, tend to report higher vegetable consumption. As weight progresses toward the Overweight categories, vegetable consumption decreases slightly, indicating a possible reduction in healthy dietary habits. However, at the upper end of the weight spectrum, corresponding to Obesity Type II and Obesity Type III, vegetable consumption increases again, potentially due to dietary interventions or awareness in this group.

The trend reveals two possible key insights:

-   A dip in vegetable consumption occurs in intermediate weight ranges, aligning with the overweight population.
-   The sharp increase in vegetable consumption among the most obese individuals may reflect lifestyle adjustments prompted by health concerns or medical advice.

##### Physical activity

Plot histogram and density.

```{r}
ggplot(dataset, aes(x = physical_act)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "skyblue", color = "black", alpha = 0.6) +
  geom_density(color = "darkblue", size = 1) +
  ggtitle("Histogram and Density of Physical Activity") +
  theme_minimal() +
  labs(x = "Physical Activity", y = "Density")
```

The histogram and density plot reveal that physical activity levels have distinct peaks at 0, 1, 2, and 3, suggesting that these values are common reported levels. Intermediate values, likely due to synthetic data or SMOTE, are also present but less frequent.

Violin plot by category.

```{r}
ggplot(dataset, aes(x = obesity_lev, y = physical_act, fill = obesity_lev)) +  # Replace 'obesity_lev' with any category variable
  geom_violin(trim = FALSE) +
  geom_boxplot(width = 0.1, color = "black", fill = "white") +
  ggtitle("Violin Plot of Physical Activity by Obesity Level") +
  theme_minimal() +
  labs(x = "Obesity Level", y = "Physical Activity") +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Physical activity levels show a slight decline as obesity levels increase, particularly evident in the narrowing distributions and lower medians observed for Obesity Type II and Obesity Type III categories. In contrast, the Insufficient Weight and Normal Weight groups exhibit higher physical activity levels, as reflected by their broader and more symmetrical distributions.

The graph reveals a distinct trend: individuals in lower obesity categories engage in more physical activity compared to those in higher obesity categories. This trend suggests an inverse relationship between physical activity and obesity levels.

##### Water consumption

Plot histogram and density for water consumption.

```{r}
ggplot(dataset, aes(x = ch2o)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "skyblue", color = "black", alpha = 0.6) +
  geom_density(color = "darkblue", size = 1) +
  ggtitle("Histogram and Density of Comsumption of Water") +
  theme_minimal() +
  labs(x = "CH2O", y = "Density")
```

This histogram and density plot of daily water consumption (CH2O) shows a clear peak at 2 liters, indicating that most individuals consume around this amount. This aligns with scientific literature, which generally recommends an average daily water intake of about 2 liters for optimal health.

Violin Plot by Gender.

```{r, message=FALSE, warning=FALSE}
# Scatterplot with a LOESS trend line
ggplot(dataset, aes(x = weight, y = ch2o, color = obesity_lev)) +
    geom_point(alpha = 0.6) +
    geom_smooth(method = "loess", se = FALSE, color = "black") +
    labs(title = "Scatterplot of Weight vs Water Consumption", x = "Weight", y = "Water Consumption (ch2o)") +
    theme_minimal() +
coord_cartesian(xlim= c(35, 135))

```

The scatterplot visualizes the relationship between weight and water consumption (ch2o), categorized by obesity levels. The trend line reveals a slightly increasing pattern of water consumption as weight increases, though the relationship is relatively weak and mostly linear.

This pattern suggests that individuals with Insufficient Weight and Normal Weight categories generally report slightly lower water consumption compared to individuals in the higher weight categories, such as Obesity Type II and III. The increase in water consumption among higher weight groups could indicate attempts to adopt healthier habits or increased hydration needs due to larger body sizes. However, the relatively flat trend across most weight ranges suggests that water consumption does not vary dramatically across different weight categories, highlighting a potential area for targeted interventions to promote hydration as a component of healthy dietary behavior.

##### Technology utilization

Histogram with Density.

```{r}
ggplot(dataset, aes(x = use_tech)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "lightblue", color = "black", alpha = 0.6) +
  geom_density(color = "blue", size = 1) +
  labs(title = "Histogram and Density of Use of Technology", x = "Use of Technology", y = "Density") +
  theme_minimal()
```

Density of Use of Technology by Obesity Level.

```{r}
ggplot(dataset, aes(x = use_tech, fill = obesity_lev)) +
  geom_density(alpha = 0.5) +
  labs(title = "Density of Use of Technology by Obesity Level", x = "Use of Technology", y = "Density") +
  theme_minimal()
```

This density plot provides a perspective on the use of technology across different obesity levels. A striking feature is the sharp, dominant peak in Obesity Type III (yellow) around the value of 1. This pattern diverges notably from the smoother and more evenly distributed curves seen in other obesity categories, suggesting a unique behavioral trend in this group.

The peak indicates a strong clustering of individuals in Obesity Type III who report moderate use of technology, which may reflect consistent engagement with technology-based activities such as sedentary work, entertainment, or even health-monitoring applications. In contrast, other obesity categories, such as Obesity Type II and Overweight Level II, exhibit more balanced distributions without a single dominant peak, hinting at more varied technology usage patterns.

This observation raises interesting questions about the role of technology in shaping lifestyle behaviors in Obesity Type III individuals. It may point to a reliance on technology that correlates with a sedentary lifestyle, a known risk factor for obesity. Alternatively, it could reflect targeted interventions or habits specific to this group. 


