# Exploratory Data Analysis (EDA)


The Exploratory Data Analysis (EDA) phase of the project was designed to uncover meaningful patterns and insights while ensuring the dataset was optimized for analysis. A correlation heatmap was employed early in the process to identify and assess relationships between variables. By comparing the initial and refined versions of the heatmap, we effectively filtered out less relevant variables, allowing the analysis to focus on the most impactful features. This step not only streamlined the dataset but also enhanced its interpretability, ensuring a more targeted exploration of key patterns.

The retained variables were selected based on their strong correlations with the target variable and their potential to reveal underlying trends. The EDA process systematically examined these variables using various visualizations to uncover distributions, trends, and potential anomalies in the data.

In the initial stage, the analysis focused on Age, Height, and Weight—variables identified as most relevant through the refined heatmap. These features were prioritized due to their strong relationships with the target outcomes and their foundational role in understanding key trends within the dataset.

## Age

```{r}
age_summary <- summary(dataset$age)
age_sd <- sd(dataset$age, na.rm = TRUE)
sum_age_df <- tibble::tibble(
  Metric = c(names(age_summary), "Standard Deviation"),
  Value = round(c(age_summary, age_sd), 2)
)
kable(sum_age_df, format = "markdown", caption = "Age Variable Statistics")
```

The age variable exhibits a right-skewed distribution, with a mean of 24.3 years and a median of 22.78 years, indicating a slight asymmetry toward younger ages. The range spans from 14 to 61 years, though the majority of individuals fall within the 20–30 age group. A standard deviation of 6.35 years reflects moderate variability in age across the dataset. This predominantly young sample may introduce limitations when generalizing findings to older populations, where obesity-related factors might differ significantly.

```{r}
g5 <- ggplot(dataset, aes(x = obesity_lev, y = age, fill = obesity_lev)) +
  geom_violin(trim = FALSE, alpha = 0.6) +
  geom_boxplot(width = 0.1, color = "black", fill = "white") +
  labs(title = "Age Distribution by Obesity Level", x = "Obesity Level", y = "Age") +
  theme_minimal() +
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
plot5 <- ggplotly(g5)
plot5
```

A violin and boxplot combination was designed to visualize the age distribution across various obesity levels, offering a detailed perspective on patterns and trends within the data. This visualization highlights that individuals with insufficient or normal weight are predominantly younger, with ages concentrated between 14 and 30 years. 
In contrast, higher obesity levels, such as Obesity Type I and Type II, display a broader age range, peaking around 30–40 years. Severe obesity (Type III) is rare among younger individuals but becomes more prevalent in mid-adulthood. This visualization emphasizes the gradual increase in obesity risk with age, underlining the importance of early intervention, particularly during early and mid-adulthood, when such risks are most pronounced.

```{r, message=FALSE, warning=FALSE}
g6 <- ggplot(dataset, aes(x = age, y = as.numeric(obesity_lev))) +
  geom_jitter(alpha = 0.3) +
  geom_smooth(method = "loess", se = FALSE, color = "blue") +
  labs(title = "Trend of Obesity Level with Age", x = "Age", y = "Obesity Level") +
  theme_minimal()
plot6 <- ggplotly(g6)
plot6
```

Complementing this, the trend line graph further captures the trajectory of obesity levels with age. A sharp rise in obesity is observed from adolescence to early adulthood, peaking in the 25–30 years range. This critical transition phase is likely influenced by lifestyle factors such as reduced physical activity, increased caloric intake, and metabolic changes. After this peak, the trend reveals a gradual decline in obesity levels beyond 30 years, potentially reflecting improved health awareness, dietary adjustments, or a selection bias in older populations. These insights underscore the mid-20s to early-30s as a pivotal stage for targeted obesity prevention and intervention strategies.

## Height

```{r}
height_summary <- summary(dataset$height)
height_sd <- sd(dataset$height, na.rm = TRUE)
sum_height_df <- tibble::tibble(
  Metric = c(names(height_summary), "Standard Deviation"),
  Value = round(c(height_summary, height_sd), 2)
)
kable(sum_height_df, format = "markdown", caption = "Height Variable Statistics")

g7 <- ggplot(dataset, aes(x = height)) +
  geom_histogram(bins = 20, fill = "purple", color = "black", alpha = 0.7) +
  labs(title = "Height Distribution", x = "Height (m)", y = "Count") +
  theme_minimal()
plot7 <- ggplotly(g7)
plot7
```

A bar chart was generated to visualize the height distribution, revealing an approximately normal shape with a slight right skew. Most values range between 1.45m and 1.98m, peaking around 1.75m, which represents the most common height. Both the mean and median are 1.7m, confirming a nearly symmetrical distribution. The standard deviation of 0.09 indicates low variability, and no extreme outliers are observed, highlighting a realistic and consistent dataset for height.

```{r}
g8 <- ggplot(dataset, aes(x = obesity_lev, y = height, fill = obesity_lev)) +
  geom_violin(alpha = 0.6) +
  labs(title = "Height Distribution by Obesity Level", x = "Obesity Level", y = "Height") +
  theme_minimal() +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))
plot8 <- ggplotly(g8)
plot8
```

The violin plot was utilized to further explores the height distribution across obesity levels. Each category exhibits relatively low variability, with overlapping ranges across groups. Insufficient and Normal Weight categories have slightly narrower distributions, centered around 1.7m. As obesity levels increase, from Obesity Type I to Type III, the distributions remain consistent, indicating that height does not significantly vary with obesity classification. These findings suggest that while height remains a stable feature, weight likely plays a more decisive role in determining obesity levels.

## Weight

```{r}
g9 <- ggplot(dataset, aes(x = weight, fill = gender)) +
  geom_density(alpha = 0.5) +
  labs(title = "Density Plot of Weight by Gender", x = "Weight", y = "Density") +
  scale_fill_manual(values = c("pink", "lightblue"), name = "Gender", labels = c("Female", "Male")) +
  theme_minimal()
plot9 <- ggplotly(g9)
plot9
```

The development of the density plot showed distinct differences in weight distribution between genders. Females generally exhibit lower weights, with a peak around 70 units, whereas males show peaks at 85 and 115 units, reflecting a tendency toward higher weights. An overlapping region between 80 and 90 units indicates common weight ranges for both genders, though the distinct peaks underscore gender-based differences. 

```{r, message=FALSE, warning=FALSE}
ggplot(dataset, aes(x = weight, y = obesity_lev, fill = obesity_lev)) +
  geom_density_ridges(scale = 0.9, alpha = 0.6) +
  labs(title = "Ridgeline Plot of Weight by Obesity Level", x = "Weight", y = "Obesity Level") +
  theme_minimal() +
  theme(legend.position = "none")

```

Building on these insights, the ridgeline plot further illustrates the relationship between weight and obesity levels. As obesity levels increase, the weight distribution shifts consistently toward higher values. Categories such as "Insufficient Weight" and "Normal Weight" cluster at lower ranges, while higher obesity types (I, II, and III) peak at significantly greater weights. 
This clear progression confirms a strong positive association between weight and obesity levels, reinforcing the centrality of weight in obesity classification. The dataset's average weight remains at 86.6 units with a standard deviation of 26.6, capturing the variability across different obesity categories.

## Height and Weight

To explore the relationship between height and weight more thoroughly, scatter plots were utilized to analyze trends across obesity levels, offering valuable insights that align with theoretical expectations and highlight weight variations within height ranges for different obesity classifications.

```{r, message=FALSE, warning=FALSE}
g11 <- ggplot(dataset, aes(x = height, y = weight, color = obesity_lev)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, aes(group = obesity_lev)) +  # Adds a trend line for each obesity level
  ggtitle("Scatter Plot of Weight vs Height by Obesity Level") +
  theme_minimal() +
  labs(x = "Height", y = "Weight", color = "Obesity Level")
plot11 <- ggplotly(g11)
plot11

g12 <- ggplot(dataset, aes(x = height, y = weight)) +
  geom_point(alpha = 0.7, aes(color = obesity_lev)) +
  facet_wrap(~ obesity_lev) +
  ggtitle("Facet Grid of Weight and Height by Obesity Level") +
  theme_minimal() +
  labs(x = "Height", y = "Weight", color = "Obesity Level") +
  theme(legend.position = "none")
plot12 <- ggplotly(g12)
plot12

```

Expanding on this analysis, the first scatter plot provides a broad overview of the positive trend between height and weight across obesity levels. While this visualization captures the general relationship, the overlapping data points obscure category-specific details, making it difficult to differentiate unique patterns within each obesity group.

To address this limitation, the data was refined into a facet grid that isolates each obesity category, offering a more nuanced perspective on weight-height trends. The facet grid reveals notable differences in the slopes of the relationships: higher obesity levels, such as Obesity Type II and III, display steeper slopes, suggesting a stronger height-weight correlation, while lower categories like Insufficient Weight and Normal Weight exhibit flatter slopes, indicating a weaker association. This detailed examination underscores the variability in the height-weight relationship across obesity levels, providing a clearer understanding of how these factors interact within distinct classifications.

```{r}
correlation_height_weight <- cor(dataset$height, dataset$weight, use = "complete.obs")
correlation_height_weight
```

The observed correlation between height and weight (r = 0.457) aligns with findings in existing literature, confirming a moderate positive relationship and reinforcing the expectation that taller individuals generally weigh more, though the strength of this association varies slightly across obesity levels.

With the analysis of age, height, and weight completed, attention shifts to exploring the remaining variables in the dataset. These variables, while less directly correlated with the target outcomes, offer critical insights into behavioral, lifestyle, and environmental factors that may influence obesity levels. 

## Food between meals

```{r}
g13 <- ggplot(dataset, aes(x = food_btw_meals, fill = obesity_lev)) +
   geom_bar(position = "dodge", color = "black") +
   ggtitle("Dodged Bar Chart for Food Between Meals by Obesity Levels") +
   labs(x = "Food Between Meals", y = "Count", fill = "Obesity Levels") +
   theme_minimal() +
   theme(
         plot.title = element_text(hjust = 0.5, size = 14))

plot13 <- ggplotly(g13)
plot13

g14 <- ggplot(dataset, aes(x = obesity_lev, fill = food_btw_meals)) +
    geom_bar(position = "fill") + # Stacked bar chart with proportions
    scale_y_continuous(labels = scales::percent_format(accuracy = 1)) + # Format y-axis as percentages
    ggtitle("Proportion of Food Between Meals Across Obesity Levels") + # Shortened and clear title
    labs(x = "Obesity Levels", y = "Proportion (%)", fill = "Food Between Meals") + # Correct axis and legend labels
    theme_minimal() +
    theme(
        axis.text.x = element_text(angle = 45, hjust = 1), # Rotate x-axis text for readability
        plot.title = element_text(hjust = 0.5, size = 14) # Center and style the title
    )
plot14 <- ggplotly(g14)
plot14
```

Together, these visualizations were designed to analyze the frequency of eating between meals, providing insights into both total prevalence and shifting trends across obesity categories.
The most dominant behavior across all categories is "Sometimes," which peaks in intermediate levels like Normal Weight and Overweight Level I, reflecting a common pattern of moderate snacking. However, as obesity levels increase to Obesity Types I–III, the responses for "Frequently" and "Always" diminish, while "Sometimes" becomes even more prevalent. This shift could indicate that higher obesity levels are more associated with habitual moderate snacking rather than excessive meal-snacking frequency. On the other hand, "No" responses remain negligible across all obesity levels, suggesting that eating between meals is almost universal in this population. 
This pattern underscores the importance of examining not just the frequency but also the quality and context of snacking as potential contributors to obesity progression.

## High-caloric food consumption

```{r}

g16 <- ggplot(dataset, aes(x = obesity_lev, fill = caloric_food)) +
  geom_bar(
    position = "dodge",
    color = "black"
  ) +
  ggtitle("Grouped Bar Chart of High-Caloric Food Consumption Across Obesity Levels") +
  labs(x = "Obesity Levels", y = "Count", fill = "High-Caloric Food Consumption") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5, size = 14)
  )
plot16 <- ggplotly(g16)
plot16

```

To investigate the consumption of high-caloric food a grouped bar chart used, highlighting a clear trend of increased high-caloric food consumption as obesity levels rise. High-caloric food consumption ("yes") is the dominant behavior across all categories, with the highest counts observed in higher obesity levels (Obesity Type I–III). In contrast, "no" responses are more visible in lower categories such as Insufficient Weight and Normal Weight, though they remain significantly lower in count compared to "yes" responses. This pattern underscores the strong association between high-caloric food intake and obesity progression.

```{r}
percentage_high_caloric_consumers <- mean(dataset$caloric_food == "yes") * 100
percentage_high_caloric_consumers
```

More precisely, a notable 88.4% of participants report frequent consumption of high-calorie foods, a behavior strongly associated with weight gain. This underscores the critical importance of dietary interventions aimed at reducing high-calorie intake to address obesity progression effectively.

## Alcohol consumption

```{r}
# Filter out "Always" responses from the dataset
filtered_dataset <- dataset %>%
  filter(freq_alcohol != "Always")

g17 <- ggplot(filtered_dataset, aes(x = freq_alcohol, fill = obesity_lev)) +
   geom_bar(position = "dodge", color = "black") +
   ggtitle("Dodged Bar Chart for Alcohol Consumption by Obesity Levels") +
   labs(x = "Alcohol Consumption Frequency", y = "Count", fill = "Obesity Levels") +
   theme_minimal() +
   theme(
         plot.title = element_text(hjust = 0.5, size = 14)) # Center and style the title

plot17 <- ggplotly(g17)
plot17
```

Regarding alcohol consumption, the chart chosen excludes "Always" responses due to their near absence in the dataset, highlighting that excessive alcohol consumption is rare. Instead, the data reveals that "Sometimes" is the dominant alcohol consumption frequency across all obesity levels, particularly in Normal Weight, Overweight Level I, and II categories. As obesity increases, "Frequently" becomes slightly more prominent, especially in Obesity Type III, while "No" responses decrease, being more common in lower obesity levels such as Insufficient and Normal Weight. This trend underlines the potential relationship between moderate-to-frequent alcohol consumption and higher obesity levels, emphasizing its importance for obesity-related behavioral research.

```{r}
# Prepare the data summary for 'Sometimes' and 'No' responses
data_summary <- dataset %>%
  filter(freq_alcohol %in% c("Sometimes", "No")) %>%
  group_by(obesity_lev, freq_alcohol) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(obesity_lev) %>%
  mutate(
    total = sum(count),
    proportion = count / total
  ) %>%
  ungroup()

# Visualization with updated title
g18 <- ggplot(data_summary, aes(x = obesity_lev, y = proportion, group = freq_alcohol, color = freq_alcohol)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +  # Format y-axis as percentages
  ggtitle("Proportion of 'Sometimes' and 'No' Alcohol Responses by Obesity Level") +
  labs(x = "Obesity Level", y = "Proportion (%)", color = "Alcohol Frequency") +
  scale_color_manual(values = c("No" = "purple", "Sometimes" = "gold")) + # Improved color scheme
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5, size = 14),  # Center and style title
    legend.position = "top"
  )
plot18 <- ggplotly(g18)
plot18
```

To better illustrate the trends in alcohol consumption frequency across obesity levels, this graph was created to highlight the shifting proportions of individuals consuming alcohol "Sometimes" and abstaining ("No"). The proportion of individuals who drink alcohol "Sometimes" shows a steady increase with higher obesity levels, peaking in Obesity_Type_III. Conversely, the proportion of those who abstain from alcohol decreases as obesity levels rise, suggesting an inverse relationship between abstention and obesity severity.

This pattern raises questions about the potential interaction between alcohol consumption frequency and caloric food preferences, as both behaviors appear to be associated with higher obesity levels. Investigating this interaction could provide insights into whether a combination of moderate alcohol consumption and high-calorie food preferences exerts a compounded effect on obesity risk. Understanding these combined lifestyle factors could inform strategies aimed at mitigating obesity progression more effectively.

## Daily Calorie Monitoring

```{r}


# Dodged Bar Chart for calorie_check by Obesity Levels
g19 <- ggplot(dataset, aes(x = calorie_check, fill = obesity_lev)) +
   geom_bar(position = "dodge", color = "black") +
   ggtitle("    Dodged Bar Chart for the check of the calories by Obesity Levels") +
   labs(x = "Check of calories", y = "Count", fill = "Obesity Levels") +
   theme_minimal() +
   theme(
         plot.title = element_text(hjust = 0.5, size = 14)) # Center and style the title
plot19 <- ggplotly(g19)
plot19

data_summary <- dataset %>%
  group_by(obesity_lev, calorie_check) %>%
  summarise(count = n(), .groups = "drop") %>%
  mutate(total = sum(count), proportion = count / total)

# Proportion of Calorie Checking by Obesity Level
g20 <- ggplot(data_summary, aes(x = obesity_lev, y = proportion, group = calorie_check, color = calorie_check)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  scale_y_continuous(labels = scales::percent) +
  scale_color_manual(values = c("no" = "lightcoral", "yes" = "lightblue")) +
  labs(title = "Proportion of Calorie Checking by Obesity Level", x = "Obesity Level", y = "Proportion", color = "Calorie Check") +
  theme_minimal() +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))
plot20 <- ggplotly(g20)
plot20

```

For the variable calorie_check, a Dodged Bar Chart was rendered, highlighting two main trends regarding calorie-checking behavior across obesity levels: a significant increase in "Yes" responses as obesity levels rise, particularly from Overweight Level II onward, and a decrease in "No" responses, which are more prevalent in lower obesity levels like Normal Weight and Insufficient Weight. 

The proportion graph simplifies these trends by clearly illustrating the proportional shift between "Yes" and "No" responses, making the contrast between lower and higher obesity levels more visually apparent. Together, these visualizations emphasize a potential association between obesity severity and an increased tendency to check calorie intake, suggesting heightened dietary awareness in higher obesity categories.

## Vegetable consumption

```{r}
g21 <- ggplot(dataset, aes(x = vegetable_food)) +
  geom_histogram(aes(y =after_stat(density)), bins = 30, fill = "lightgreen", color = "black", alpha = 0.6) +
  geom_density(color = "darkgreen", linewidth = 1) +
  ggtitle("Histogram and Density of Vegetable Food Consumption") +
  theme_minimal() +
  labs(x = "Vegetable Food Consumption", y = "Density")
plot21 <- ggplotly(g21)
plot21

```

```{r, message=FALSE, warning=FALSE}
g22 <- ggplot(dataset, aes(x = weight, y = vegetable_food, color = obesity_lev)) +
    geom_point(alpha = 0.6) +
    geom_smooth(method = "loess", se = FALSE, color = "black") +
    labs(title = "Scatterplot of Weight vs Vegetable Food Consumption", 
         x = "Weight", 
         y = "Vegetable Food Consumption") +
    theme_minimal() +
    coord_cartesian(xlim= c(40, 135), ylim= c(2, 3))
plot22 <- ggplotly(g22)
plot22
```

To explore the relationship between vegetable food consumption and weight, two visualizations were employed. A histogram combined with a density plot was used to examine the overall distribution of vegetable consumption across the dataset. This provided insights into the frequency and variability of vegetable intake levels, with distinct peaks indicating common consumption behaviors. Following this, a scatterplot was generated with a trend line illustrating a distinct, non-linear relationship: vegetable consumption initially decreases as weight increases but then begins to rise again at higher weight levels.

This pattern suggests that individuals with lower weight, particularly those in the Insufficient Weight and Normal Weight categories, tend to report higher vegetable consumption. As weight progresses toward the Overweight categories, vegetable consumption decreases slightly, indicating a possible reduction in healthy dietary habits. However, at the upper end of the weight spectrum, corresponding to Obesity Type II and Obesity Type III, vegetable consumption increases again, potentially due to dietary interventions or awareness in this group.

The trend reveals two possible key insights:

-   A dip in vegetable consumption occurs in intermediate weight ranges, aligning with the overweight population.
-   The sharp increase in vegetable consumption among the most obese individuals may reflect lifestyle adjustments prompted by health concerns or medical advice.

## Physical activity

To explore how physical activity varies across obesity levels, two complementary visualizations were utilized.

```{r, warning=FALSE}
g23 <- ggplot(dataset, aes(x = obesity_lev, y = physical_act, fill = obesity_lev)) +  
  geom_violin(trim = FALSE, alpha = 0.6) +
  geom_boxplot(width = 0.1, color = "black", fill = "white") +
    ggtitle("Violin Plot of Physical Activity by Obesity Level") +
  labs(x = "Obesity Level", y = "Physical Activity") +
  theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))
plot23 <- ggplotly(g23)
plot23


# Calculate the average physical activity per obesity level
average_activity <- dataset %>%
  group_by(obesity_lev) %>%
  summarise(mean_activity = mean(physical_act))

# Plot the average physical activity
g24 <- ggplot(average_activity, aes(x = obesity_lev, y = mean_activity, group = 1)) +
  geom_line(color = "blue", size = 1.2) +
  geom_point(size = 3, color = "darkblue") +
  ggtitle("Average Physical Activity Across Obesity Levels") +
  labs(x = "Obesity Level", y = "Average Physical Activity") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5, size = 14))
plot24 <- ggplotly(g24)
plot24

```

The line chart emphasizes the consistent decline in average physical activity as obesity levels increase. Starting from Insufficient and Normal Weight categories, the average activity levels gradually drop, stabilizing briefly at Obesity Types I and II before sharply decreasing in Obesity Type III. This trend underscores a clear negative association between physical activity and obesity levels, highlighting the potential role of physical inactivity as a contributing factor to severe obesity. Combined, these visualizations illustrate the importance of promoting physical activity as a preventive and intervention strategy across all weight categories, particularly for individuals at higher obesity levels.

## Water consumption

```{r, warning=FALSE}
g25 <- ggplot(dataset, aes(x = ch2o)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "skyblue", color = "black", alpha = 0.6) +
  geom_density(color = "darkblue", size = 1) +
  ggtitle("Histogram and Density of Comsumption of Water") +
  theme_minimal() +
  labs(x = "CH2O", y = "Density")
plot25 <- ggplotly(g25)
plot25
```

About the distribution of daily water consumption (CH2O) an histogram with density plot chart was generated to show a clear peak at 2 liters, indicating that most individuals consume around this amount. This aligns with scientific literature, which generally recommends an average daily water intake of about 2 liters for optimal health.

```{r, message=FALSE}
g26 <- ggplot(dataset, aes(x = weight, y = ch2o)) +
    geom_point(alpha = 0.5, color = "lightgray") +  # Scatterplot in the background with reduced opacity
    geom_smooth(method = "loess", se = FALSE, color = "blue", linewidth = 1.2) +  # Trend line in the foreground
    labs(title = "Scatterplot with Trend Line: Water Consumption vs Weight", 
         x = "Weight", y = "Water Consumption (ch2o)") +
    theme_minimal()
plot26 <- ggplotly(g26)
plot26
```

To explore the connection further, a scatterplot with a trend line was created to analyze how water intake varies with weight. The trend line indicates a slight upward pattern, suggesting that water consumption increases moderately as weight rises. Individuals in lower weight categories, such as Insufficient and Normal Weight, show slightly lower water consumption compared to those in higher weight groups, including Obesity Type II and III. This increase in consumption among heavier individuals could reflect greater hydration needs or attempts to implement healthier habits. Despite this trend, water intake appears relatively stable across most weight ranges, highlighting an opportunity to encourage better hydration practices as part of a comprehensive health approach.

## Technology utilization

```{r}
g28 <- ggplot(dataset, aes(x = use_tech, fill = obesity_lev)) +
  geom_density(alpha = 0.5) +
  labs(title = "Density of Use of Technology by Obesity Level", x = "Use of Technology", y = "Density") +
  theme_minimal()
plot28 <- ggplotly(g28)
plot28
```

This density plot provides a perspective on the use of technology across different obesity levels. A striking feature is the sharp, dominant peak in Obesity Type III (yellow) around the value of 1. This pattern diverges notably from the smoother and more evenly distributed curves seen in other obesity categories, suggesting a unique behavioral trend in this group.

The peak indicates a strong clustering of individuals in Obesity Type III who report moderate use of technology, which may reflect consistent engagement with technology-based activities such as sedentary work, entertainment, or even health-monitoring applications. In contrast, other obesity categories, such as Obesity Type II and Overweight Level II, exhibit more balanced distributions without a single dominant peak, hinting at more varied technology usage patterns.

This observation raises interesting questions about the role of technology in shaping lifestyle behaviors in Obesity Type III individuals. It may point to a reliance on technology that correlates with a sedentary lifestyle, a known risk factor for obesity. Alternatively, it could reflect targeted interventions or habits specific to this group.

The Exploratory Data Analysis (EDA) phase provided a comprehensive understanding of the dataset, offering key insights into the relationships between various behavioral, lifestyle, and demographic factors and obesity levels. By focusing on critical variables the EDA revealed patterns and trends that are integral to the modeling process.



