
# 3. Exploratory Data Analysis (EDA)

#### 3.1 Descriptive statistics and distribution analysis

##### Age

Descriptive statistic for Age

```{r}
summary(dataset$age)
sd(dataset$age, na.rm = TRUE)
```

Age distribution.

```{r}
ggplot(dataset, aes(x = age)) +
  geom_histogram(bins = 20, fill = "skyblue", color = "black", alpha = 0.7) +
  labs(title = "Age Distribution", x = "Age", y = "Count")+
  theme_minimal()
```

The average age of participants is 24.3 years, with a standard deviation of 6.35 years, predominantly concentrated in the 20-30 age range This distribution suggests a young population, which may limit the generalizability of results to older age groups, where risk factors for overweight and obesity may differ.

Age by Obesity Level

Distribution of Age by Obesity Level.

```{r}
ggplot(dataset, aes(x = age, fill = obesity_lev)) +
  geom_histogram(bins = 20, color = "black", alpha = 0.6) +
  facet_wrap(~ obesity_lev) +
  ggtitle("Distribution of Age by Obesity Level") +
  labs(x = "Age", y = "Count") +
  theme_minimal()
```

The age distribution varies across obesity levels, with younger ages being more prevalent in lower obesity levels (like Insufficient Weight and Normal Weight), while higher obesity levels tend to have older individuals. This suggests a possible trend where age might correlate with obesity level, especially in higher obesity categories.

Age Distribution by Obesity Level (Violin Plot).

```{r}
ggplot(dataset, aes(x = obesity_lev, y = age, fill = obesity_lev)) +
  geom_violin(trim = FALSE, alpha = 0.6) +
  geom_boxplot(width = 0.1, color = "black", fill = "white") +
  labs(title = "Age Distribution by Obesity Level", x = "Obesity Level", y = "Age") +
  theme_minimal()
```

The violin plot shows, more clearly, the spread and density of ages across obesity levels. Younger ages dominate in the lower obesity levels, while there is a wider age range with a higher density around 30–40 years in the higher obesity levels. This pattern supports the idea that older age groups are more likely to fall into higher obesity categories.

Age Distribution with SMOOTH Trend Line for Obesity Probability.

```{r}
ggplot(dataset, aes(x = age, y = as.numeric(obesity_lev))) +
  geom_jitter(alpha = 0.3) +
  geom_smooth(method = "loess", se = FALSE, color = "blue") +
  labs(title = "Trend of Obesity Level with Age", x = "Age", y = "Obesity Level") +
  theme_minimal()
```

The trend line suggests an increase in obesity level with age until around 30–35, followed by a decrease. This implies that middle age might be a peak period for higher obesity levels, and there may be a trend of reducing obesity levels at older ages.

##### Height

Descriptive statistic for Height.

```{r}
summary(dataset$height)
sd(dataset$height, na.rm = TRUE)
```

Height distribution.

```{r}
ggplot(dataset, aes(x = height)) +
  geom_histogram(bins = 20, fill = "purple", color = "black", alpha = 0.7) +
  labs(title = "Height Distribution", x = "Height (m)", y = "Count") +
  theme_minimal()
```

Height by gender

Density plot for height distribution by gender.

```{r}
ggplot(dataset, aes(x = height, fill = gender)) +
  geom_density(alpha = 0.5) +
  labs(title = "Density Plot of Height by Gender", x = "Height", y = "Density") +
  scale_fill_manual(values = c("pink", "lightblue"), name = "Gender", labels = c("Female", "Male")) +
  theme_minimal()
```

The average height is 1.70 m (SD = 0.09 m), with a notable difference between genders, where males tend to have a higher median height.

Height by Obesity Level

Box Plot of Height by Obesity Level.

```{r}
ggplot(dataset, aes(x = obesity_lev, y = height, fill = obesity_lev)) +
  geom_boxplot(alpha = 0.6) +
  labs(title = "Height Distribution by Obesity Level", x = "Obesity Level", y = "Height") +
  theme_minimal() +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))
```

The boxplot indicates that individuals with lower obesity levels (e.g., Insufficient Weight, Normal Weight) tend to have a more consistent height range, while those in higher obesity levels (like Obesity Type II and III) show more variability in height. This suggests that weight may be more influential than height alone in determining obesity level.

##### Weight

Descriptive statistic for Weight.

```{r}
summary(dataset$weight)
sd(dataset$weight, na.rm = TRUE)
```

Weight by gender

Density plot for weight distribution by gender.

```{r}
ggplot(dataset, aes(x = weight, fill = gender)) +
  geom_density(alpha = 0.5) +
  labs(title = "Density Plot of Weight by Gender", x = "Weight", y = "Density") +
  scale_fill_manual(values = c("pink", "lightblue"), name = "Gender", labels = c("Female", "Male")) +
  theme_minimal()
```

The density plot reveals distinct weight distributions between genders. Females generally weigh less, with a peak around 60-70 units, while males peak around 90-100 and 110-120 units, indicating a tendency toward higher weights. The overlapping region around 80-90 units shows weights common to both genders, though males dominate at higher ranges Weight ranges from 39 to 173 units, with an average (mean) weight of 86.6 units. The median weight is 83 units, with a standard deviation of 26.2, indicating moderate spread.

Weight by obesity level

Ridgeline Plot of Weight by Obesity Level.

```{r}
ggplot(dataset, aes(x = weight, y = obesity_lev, fill = obesity_lev)) +
  geom_density_ridges(scale = 0.9, alpha = 0.6) +
  labs(title = "Ridgeline Plot of Weight by Obesity Level", x = "Weight", y = "Obesity Level") +
  theme_minimal() +
  theme(legend.position = "none")
```

This ridgeline plot shows a clear progression in weight distribution across different obesity levels. As the obesity level increases, the weight distribution shifts progressively to higher ranges. "Normal Weight" and "Insufficient Weight" categories are concentrated at lower weights, while higher obesity types (I, II, and III) peak at significantly greater weights, indicating a strong positive association between weight and obesity level The weight distribution has an average of 86.6 kg and a standard deviation of 26.6 kg.

##### Height and Weight

Scatter Plot (height vs weight), colored by obesity level.

```{r}
ggplot(dataset, aes(x = height, y = weight, color = obesity_lev)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, aes(group = obesity_lev)) +  # Adds a trend line for each obesity level
  ggtitle("Scatter Plot of Weight vs Height by Obesity Level") +
  theme_minimal() +
  labs(x = "Height", y = "Weight", color = "Obesity Level")
```

Facet Grid for Height and Weight by Obesity Level.

```{r}
ggplot(dataset, aes(x = height, y = weight)) +
  geom_point(alpha = 0.7, aes(color = obesity_lev)) +
  facet_wrap(~ obesity_lev) +
  ggtitle("Facet Grid of Weight and Height by Obesity Level") +
  theme_minimal() +
  labs(x = "Height", y = "Weight", color = "Obesity Level") +
  theme(legend.position = "none")
```

The scatter plot with trend lines for each obesity level reveals a clear positive correlation between weight and height across all obesity levels. As the obesity level increases, the slope generally becomes steeper, indicating a stronger weight gain relative to height. We created the facet grid to show more clearly the trends to show more clearly how The "Obesity_Type_III" (yellow) category has the steepest slope, suggesting a significant weight increase per unit of height, which is consistent with the highest level of obesity.

Correlation between height and weight.

```{r}
correlation_height_weight <- cor(dataset$height, dataset$weight, use = "complete.obs")
correlation_height_weight
```

The correlation observed between height and weight (r = 0.463) aligns with existing literature, confirming the expected positive relationship between these variables.

##### Function for EDA on categorical variables

Function for EDA on categorical variables, with separate distributions for gender and obesity levels.

```{r}
eda_categorical <- function(data, variable, gender_var = "gender", obesity_var = "obesity_lev") {
  cat("\n--- EDA for Categorical Variable:", variable, "---\n")
# Frequency and Proportion Tables
cat("\nFrequency Table:\n")
freq_table <- table(data[[variable]])
print(freq_table)


cat("\nProportion Table (Rounded):\n")
prop_table <- round(prop.table(freq_table), 2)
print(prop_table)
 
# Bar Chart with Counts (distribution)
p1 <- ggplot(data, aes(x = .data[[variable]])) +
  geom_bar(fill = "skyblue", color = "black") +
  ggtitle(paste("Distribution of", variable, "- Counts")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(y = "Count")
  print(p1)

# Distribution by Gender
p2 <- ggplot(data, aes(x = .data[[variable]], fill = .data[[gender_var]])) +
   geom_bar(position = "dodge", color = "black") +
   ggtitle(paste("Distribution of", variable, "by Gender")) +
   labs(fill = "Gender") +
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
   labs(y = "Count")
   print(p2)
  
# Dodged Bar Chart for Categorical Variable by Obesity Levels
p3 <- ggplot(data, aes(x = .data[[variable]], fill = .data[[obesity_var]])) +
   geom_bar(position = "dodge", color = "black") +
   ggtitle(paste("Dodged Bar Chart of", variable, "by", obesity_var)) +
   labs(y = "Count", fill = obesity_var) +
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
   print(p3)
  
# Stacked Bar Chart of Food Between Meals by Obesity Level (Proportions within each Obesity Level)
p4 <- ggplot(data, aes(x = .data[[obesity_var]], fill = .data[[variable]])) +
   geom_bar(position = "fill") +
   scale_y_continuous(labels = scales::percent) +
   ggtitle(paste("Stacked Bar Chart of", variable, "by", obesity_var, "- Proportions within each Obesity Level")) +
   labs(x = obesity_var, y = "Proportion", fill = variable) +
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
   print(p4)
}

#Food between meals (snacking) 
eda_categorical(dataset, "food_btw_meals", gender_var = "gender", obesity_var = "obesity_lev")
```

Overall distribution - There are very few who selected "Always" or "no."

Distribution by Gender - it appears that a slightly higher proportion of males selected "Sometimes" compared to females.

Distribution by Obesity Level - As obesity levels increase, there is a noticeable shift towards "Sometimes" and "Frequently" responses, particularly in higher obesity levels (Type I, II, and III).

Individuals with "Insufficient Weight" and "Normal Weight" show a relatively balanced spread between "no," "Sometimes," and "Frequently."

These observations suggest a potential correlation between the frequency of eating between meals and obesity levels. Individuals who report eating between meals "Sometimes" or "Frequently" seem more likely to have higher obesity levels.

In our further work we can explore the interaction with people who engage in regular physical activity, caloric_food often have increased caloric needs and may snack more frequently. Also ch20, people who drink less water may have more cravings or perceive thirst as hunger, leading them to snack more.

##### High-calore food consumption

```{r}
eda_categorical(dataset, "caloric_food", gender_var = "gender", obesity_var = "obesity_lev")
```

Overall distribution - The majority of individuals prefer caloric food, as indicated by the high count of "yes" responses.

Distribution by Gender - Both males and females show a similar pattern, with a strong preference for caloric food ("yes"), although the preference is slightly higher in males.

Distribution by Obesity Level - As obesity levels increase, the preference for caloric food ("yes") also increases, particularly in higher obesity categories (Obesity Types I, II, and III).

Proportion Analysis by Obesity Level - Individuals with higher obesity levels are overwhelmingly more likely to prefer caloric food, with nearly 100% of those in Obesity Types II and III choosing "yes".

Count participants who answered "yes" for frequent high-calorie food consumption.

```{r}
num_high_caloric_consumers <- sum(dataset$caloric_food == "yes")
```

Calculate the total number of participants in the dataset.

```{r}
total_participants <- nrow(dataset)
```

Calculate the percentage of high-calorie consumers.

```{r}
percentage_high_caloric_consumers <- (num_high_caloric_consumers / total_participants) * 100
percentage_high_caloric_consumers
```

A notable 88.4% of participants report frequent consumption of high-calorie foods, which may contribute to weight gain. This trend highlights the need for dietary interventions focused on reducing high-calorie intake.

##### Alcohol consumption

Frequence in consumption of alcohol.

```{r}
eda_categorical(dataset, "freq_alcohol", gender_var = "gender", obesity_var = "obesity_lev")
```

The majority of individuals report drinking alcohol "Sometimes" or "no," with very few indicating "Frequently" or "Always.".

Distribution by Gender - Both males and females primarily drink "Sometimes" or "no," with females showing a slightly higher preference for "Sometimes".

Distribution by Obesity Level - In lower obesity levels (Insufficient and Normal Weight), a significant proportion reports "no" alcohol consumption,In higher obesity levels, particularly Obesity_Type_I and above, there is an increased tendency to drink "Sometimes," while "no" responses decline.

Proportion Analysis by Obesity Level - Individuals with higher obesity levels are overwhelmingly more likely to prefer caloric food, with nearly 100% of those in Obesity Types II and III choosing "yes".

Line plot to show the pattern.

```{r}
# Prepare the data summary for 'Sometimes' and 'No' responses
data_summary <- dataset %>%
  filter(freq_alcohol %in% c("Sometimes", "No")) %>%
  group_by(obesity_lev, freq_alcohol) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(obesity_lev) %>%
  mutate(total = sum(count), proportion = count / total) %>%
  ungroup()

# Check if data_summary has the expected columns
print(head(data_summary))

ggplot(data_summary, aes(x = obesity_lev, y = proportion, group = freq_alcohol, color = freq_alcohol)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  scale_y_continuous(labels = scales::percent) +
  ggtitle("Proportion of 'Sometimes' and 'No' Responses for Alcohol Frequency by Obesity Level") +
  labs(x = "Obesity Level", y = "Proportion", color = "Alcohol Frequency") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```

The proportion of individuals who drink alcohol "Sometimes" increases with higher obesity levels, peaking in Obesity_Type_III. In contrast, the likelihood of abstaining from alcohol ("no") decreases as obesity levels rise. This pattern suggests that moderate alcohol consumption may be associated with higher obesity levels, while abstention is more common among those with lower obesity levels.

A possible interaction to investigate later is between alcohol frequency and caloric food preference, as both behaviors seem linked to higher obesity levels. Exploring this could reveal if individuals with a preference for caloric foods and moderate alcohol consumption have a compounding effect on obesity risk. This investigation could help clarify whether combined lifestyle factors contribute more significantly to higher obesity levels than each factor alone.

Monitoring of the calories in the day.

```{r}
eda_categorical(dataset, "calorie_check", gender_var = "gender", obesity_var = "obesity_lev")
```

The vast majority of individuals do not check calories.

Distribution by Gender - Only a small percentage in each gender reporting that they check calories.

Distribution by Obesity Level - There is a slight increase in calorie-checking behavior among individuals with lower obesity levels (Insufficient and Normal Weight).

Proportion Analysis by Obesity Level - Calorie-checking behavior slightly decreases as obesity levels increase.

```{r}
data_summary <- dataset %>%
  group_by(obesity_lev, calorie_check) %>%
  summarise(count = n(), .groups = "drop") %>%
  mutate(total = sum(count), proportion = count / total)

ggplot(data_summary, aes(x = obesity_lev, y = proportion, group = calorie_check, color = calorie_check)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  scale_y_continuous(labels = scales::percent) +
  scale_color_manual(values = c("no" = "lightcoral", "yes" = "lightblue")) +
  labs(title = "Proportion of Calorie Checking by Obesity Level", x = "Obesity Level", y = "Proportion", color = "Calorie Check") +
  theme_minimal()
```

##### Vegetable consumption

Distribution of vegetable consumption frequency (vegetable_food): initially, a bar chart was created. This allowed us to clearly identify the presence of non-integer data. Values are rounded to the nearest integer, and the categories are renamed to make the informational content immediately understandable.

```{r}
dataset$vegetable_food <- round(dataset$vegetable_food)
ggplot(dataset, aes(x = factor(vegetable_food, levels = c(1, 2, 3), 
                               labels = c("Rarely", "Sometimes", "Often")))) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "Vegetable Consumption Frequency", x = "Frequency of Vegetable Consumption", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=0.5))
```

Vegetable intake is generally moderate, with a mean score of 2.4 on a 1-to-3 scale (1 = "Rarely", 3 = "Often").

However, a detailed distribution of consumption and an analysis of associations with other risk factors are necessary to make informed statements about the quality of dietary habits and potential health risks.

##### Number of meals per day

The same issues as with the previous variable were initially encountered in the graphical representation. Similarly, the values were rounded and the labels were renamed.

```{r}
ggplot(dataset, aes(x = factor(round(nb_meal_day), levels = c(1, 2, 3, 4, 5), 
                               labels = c("1 Meal", "2 Meals", "3 Meals", "4 Meals", "5+ Meals")))) + 
  geom_bar(fill = "orange", color = "black") + 
  labs(title = "Number of Meals per Day", x = "Meals per Day", y = "Count") + 
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust = 0.5))
```

Most participants consume three meals daily, with a mean of 2.7 and a standard deviation of 0.5.

Calculate the correlation between the number of meals per day and weight. Select the variables for correlation: number of meals per day and weight.

```{r}
correlation_meals_weight <- cor(dataset$nb_meal_day, dataset$weight, use = "complete.obs")
correlation_meals_weight
```

The correlation between the number of meals per day and weight is relatively weak; this suggests that meal frequency alone may not have a direct impact on weight; instead, other factors like meal quality and portion sizes could play a more significant role.

##### Physical activity

Plot histogram and density.

```{r}
ggplot(dataset, aes(x = physical_act)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "skyblue", color = "black", alpha = 0.6) +
  geom_density(color = "darkblue", size = 1) +
  ggtitle("Histogram and Density of Physical Activity") +
  theme_minimal() +
  labs(x = "Physical Activity", y = "Density")
```

The histogram and density plot reveal that physical activity levels have distinct peaks at 0, 1, 2, and 3, suggesting that these values are common reported levels. Intermediate values, likely due to synthetic data or SMOTE, are also present but less frequent.

Violin plot by category.

```{r}
ggplot(dataset, aes(x = obesity_lev, y = physical_act, fill = obesity_lev)) +  # Replace 'obesity_lev' with any category variable
  geom_violin(trim = FALSE) +
  geom_boxplot(width = 0.1, color = "black", fill = "white") +
  ggtitle("Violin Plot of Physical Activity by Obesity Level") +
  theme_minimal() +
  labs(x = "Obesity Level", y = "Physical Activity") +
  theme(legend.position = "none")
```

Individuals with lower obesity levels (e.g., Insufficient and Normal Weight) have a wider spread of physical activity levels, often skewed toward higher activity #Higher obesity levels show a tendency toward lower physical activity.

Scatter Plot with Age.

```{r}
ggplot(dataset, aes(x = age, y = physical_act)) +  # Replace 'age' with another continuous variable
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", color = "blue", se = FALSE) +
  ggtitle("Scatter Plot of Physical Activity vs Age") +
  theme_minimal() +
  labs(x = "Age", y = "Physical Activity")
```

There appears to be a negative trend between age and physical activity levels, with younger individuals tending to report higher physical activity levels.

##### Water consumption

Plot histogram and density for water consumption.

```{r}
ggplot(dataset, aes(x = ch2o)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "skyblue", color = "black", alpha = 0.6) +
  geom_density(color = "darkblue", size = 1) +
  ggtitle("Histogram and Density of Comsumption of Water") +
  theme_minimal() +
  labs(x = "CH2O", y = "Density")
```

This histogram and density plot of daily water consumption (CH2O) shows a clear peak at 2 liters, indicating that most individuals consume around this amount. This aligns with scientific literature, which generally recommends an average daily water intake of about 2 liters for optimal health.

Facet Plot (CH2O by Obesity Level and Gender).

```{r}
ggplot(dataset, aes(x = ch2o, fill = gender)) +
  geom_histogram(bins = 20, alpha = 0.6, position = "dodge") +
  facet_wrap(~ obesity_lev) +
  ggtitle("Histogram of CH2O by Obesity Level and Gender") +
  theme_minimal() +
  labs(x = "CH2O", y = "Count", fill = "Gender")
```

An uncommon trend is shown in Obesity Type II predominantly includes males, while Obesity Type III mostly includes females, highlighting a gender disparity at higher obesity levels in water consumption patterns.

Scatter Plot (ch2o vs Age).

```{r}
ggplot(dataset, aes(x = age, y = ch2o)) +
  geom_point(alpha = 0.5, color = "darkblue") +
  geom_smooth(method = "lm", color = "red", linetype = "dashed") +
  ggtitle("Scatter Plot of CH2O vs Age with Trendline") +
  theme_minimal() +
  labs(x = "Age", y = "CH2O")
```

Slight downward trend suggests a minor decrease in water consumption with age, though variability remains high across all ages.

Violin Plot by Gender.

```{r}
ggplot(dataset, aes(x = gender, y = ch2o, fill = gender)) +
  geom_violin(trim = FALSE, alpha = 0.7) +
  ggtitle("Violin Plot of CH2O by Gender") +
  theme_minimal() +
  labs(x = "Gender", y = "CH2O") +
  theme(legend.position = "none")
```

Males show more variability in water consumption compared to females, with similar median values around 2. The water consumption variable (CH2O) shows a predominant consumption level around 2 across different obesity levels, with minor variations by gender and obesity category. Originally, CH2O values were discrete (1.0, 2.0, 3.0), suggesting categorical consumption levels. However, after applying the SMOTE algorithm to address class imbalance, interpolated values emerged, resulting in a more continuous distribution. This variable's distribution across obesity levels, along with gender differences at extreme obesity levels (e.g., more males in Obesity Type II and females in Obesity Type III), indicates that CH2O could be a valuable predictor for obesity, capturing both consumption habits and subtle demographic patterns.

##### Technology utilization

Histogram with Density.

```{r}
ggplot(dataset, aes(x = use_tech)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "lightblue", color = "black", alpha = 0.6) +
  geom_density(color = "blue", size = 1) +
  labs(title = "Histogram and Density of Use of Technology", x = "Use of Technology", y = "Density") +
  theme_minimal()
```

The histogram shows a strong concentration at discrete values (0, 1, and 2), likely reflecting the original categorical nature of the data before SMOTE. This also aligns with observed density peaks.

Density of Use of Technology by Obesity Level.

```{r}
ggplot(dataset, aes(x = use_tech, fill = obesity_lev)) +
  geom_density(alpha = 0.5) +
  labs(title = "Density of Use of Technology by Obesity Level", x = "Use of Technology", y = "Density") +
  theme_minimal()
```

The density plot reveals distinct peaks in technology usage across obesity levels, with some levels like Obesity_Type_III having higher peaks, suggesting varied levels of technology use in these categories.

Boxplot by Obesity Level.

```{r}
ggplot(dataset, aes(x = obesity_lev, y = use_tech, fill = obesity_lev)) +
  geom_boxplot() +
  labs(title = "Boxplot of Use of Technology by Obesity Level", x = "Obesity Level", y = "Use of Technology") +
  theme_minimal() +
  theme(legend.position = "none")
```

Technology use shows some variation across obesity levels, with certain categories (like Obesity_Type_I and Obesity_Type_III) showing higher median usage compared to others.

Scatter Plot with Age.

```{r}
ggplot(dataset, aes(x = age, y = use_tech)) +
  geom_point(alpha = 0.4, color = "blue") +
  geom_smooth(method = "lm", color = "red", linetype = "dashed") +
  labs(title = "Scatter Plot of Use of Technology vs Age", x = "Age", y = "Use of Technology") +
  theme_minimal()
```

There is a noticeable negative correlation between age and technology use, indicating younger individuals tend to use technology more than older ones.

The 'Use of Technology' variable shows a clear trend where younger individuals use technology more frequently, as seen in the negative correlation with age. Differences in technology usage across obesity levels suggest it might have predictive value in distinguishing between levels, although the SMOTE algorithm has introduced interpolated values that blur strict categories. This variable could therefore help predict obesity levels, especially if technology usage patterns are indicative of lifestyle factors associated with obesity.
