# Exploratory Data Analysis (EDA)


The Exploratory Data Analysis (EDA) phase of the project was designed to uncover meaningful patterns and insights while ensuring the dataset was optimized for analysis. A correlation heatmap was employed early in the process to identify and assess relationships between variables. By comparing the initial and refined versions of the heatmap, we effectively filtered out less relevant variables, allowing the analysis to focus on the most impactful features. This step not only streamlined the dataset but also enhanced its interpretability, ensuring a more targeted exploration of key patterns.

Key variables retained after this step were selected based on their strong correlations with the target variable and their potential relevance to underlying patterns in the data.

The EDA process involved a systematic exploration of the cleaned and refined dataset, utilizing visualization tools to highlight trends, distributions, and potential anomalies. 

#### Descriptive statistics and distribution analysis


The initial phase of the Exploratory Data Analysis (EDA) concentrated on examining the most strongly correlated variables: Age, Height, and Weight. These variables were prioritized due to their direct relevance and significant relationships with the target outcomes, as highlighted in the refined correlation heatmap. 

##### Age

```{r}
age_summary <- summary(dataset$age)
age_sd <- sd(dataset$age, na.rm = TRUE)
sum_age_df <- tibble::tibble(
  Metric = c(names(age_summary), "Standard Deviation"),
  Value = round(c(age_summary, age_sd), 2)
)
kable(sum_age_df, format = "markdown", caption = "Age Variable Statistics")
```

The age variable exhibits a right-skewed distribution, with a mean of 24.3 years and a median of 22.78 years, indicating a slight asymmetry toward younger ages. The range spans from 14 to 61 years, though the majority of individuals fall within the 20–30 age group. A standard deviation of 6.35 years reflects moderate variability in age across the dataset. This predominantly young sample may introduce limitations when generalizing findings to older populations, where obesity-related factors might differ significantly.

Age Distribution by Obesity Level

```{r}
g5 <- ggplot(dataset, aes(x = obesity_lev, y = age, fill = obesity_lev)) +
  geom_violin(trim = FALSE, alpha = 0.6) +
  geom_boxplot(width = 0.1, color = "black", fill = "white") +
  labs(title = "Age Distribution by Obesity Level", x = "Obesity Level", y = "Age") +
  theme_minimal() +
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
plot5 <- ggplotly(g5)
plot5
```

The violin plot highlights the age distribution across obesity levels, illustrating that individuals with insufficient or normal weight are predominantly younger, with ages concentrated between 14 and 30 years. In contrast, higher obesity levels, such as Obesity Type I and Type II, display a broader age range, peaking around 30–40 years. Severe obesity (Type III) is rare among younger individuals but becomes more prevalent in mid-adulthood. This visualization emphasizes the gradual increase in obesity risk with age, underlining the importance of early intervention, particularly during early and mid-adulthood, when such risks are most pronounced.


Age Distribution with SMOOTH Trend Line for Obesity Probability

```{r, message=FALSE, warning=FALSE}
g6 <- ggplot(dataset, aes(x = age, y = as.numeric(obesity_lev))) +
  geom_jitter(alpha = 0.3) +
  geom_smooth(method = "loess", se = FALSE, color = "blue") +
  labs(title = "Trend of Obesity Level with Age", x = "Age", y = "Obesity Level") +
  theme_minimal()
plot6 <- ggplotly(g6)
plot6
```

Complementing this, the trend line graph further captures the trajectory of obesity levels with age. A sharp rise in obesity is observed from adolescence to early adulthood, peaking in the 25–30 years range. This critical transition phase is likely influenced by lifestyle factors such as reduced physical activity, increased caloric intake, and metabolic changes. After this peak, the trend reveals a gradual decline in obesity levels beyond 30 years, potentially reflecting improved health awareness, dietary adjustments, or a selection bias in older populations. These insights underscore the mid-20s to early-30s as a pivotal stage for targeted obesity prevention and intervention strategies.

##### Height


```{r}
height_summary <- summary(dataset$height)
height_sd <- sd(dataset$height, na.rm = TRUE)
sum_height_df <- tibble::tibble(
  Metric = c(names(height_summary), "Standard Deviation"),
  Value = round(c(height_summary, height_sd), 2)
)
kable(sum_height_df, format = "markdown", caption = "Height Variable Statistics")
```

Height distribution

```{r}
g7 <- ggplot(dataset, aes(x = height)) +
  geom_histogram(bins = 20, fill = "purple", color = "black", alpha = 0.7) +
  labs(title = "Height Distribution", x = "Height (m)", y = "Count") +
  theme_minimal()
plot7 <- ggplotly(g7)
plot7
```

The height distribution, as shown in the histogram, follows an approximately normal shape with a slight right skew. Most values range between 1.45m and 1.98m, peaking around 1.8m, which represents the most common height. Both the mean and median are 1.7m, confirming a nearly symmetrical distribution. The standard deviation of 0.09 indicates low variability, and no extreme outliers are observed, highlighting a realistic and consistent dataset for height.


Box Plot of Height by Obesity Level

```{r}
g8 <- ggplot(dataset, aes(x = obesity_lev, y = height, fill = obesity_lev)) +
  geom_violin(alpha = 0.6) +
  labs(title = "Height Distribution by Obesity Level", x = "Obesity Level", y = "Height") +
  theme_minimal() +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))
plot8 <- ggplotly(g8)
plot8
```

The violin plot further explores the height distribution across obesity levels. Each category exhibits relatively low variability, with overlapping ranges across groups. Insufficient and Normal Weight categories have slightly narrower distributions, centered around 1.7m. As obesity levels increase, from Obesity Type I to Type III, the distributions remain consistent, indicating that height does not significantly vary with obesity classification. These findings suggest that while height remains a stable feature, weight likely plays a more decisive role in determining obesity levels.

##### Weight

```{r}
sum_weight_df <- tibble::tibble(
  Metric = c(names(summary(dataset$weight)), "Std. Dev"),
  Value = round(c(summary(dataset$weight), sd(dataset$weight, na.rm = TRUE)), 2)
)
kable(sum_weight_df, format = "markdown", caption = "Weight Variable Statistics")
```

Density plot for weight distribution by gender

```{r}
g9 <- ggplot(dataset, aes(x = weight, fill = gender)) +
  geom_density(alpha = 0.5) +
  labs(title = "Density Plot of Weight by Gender", x = "Weight", y = "Density") +
  scale_fill_manual(values = c("pink", "lightblue"), name = "Gender", labels = c("Female", "Male")) +
  theme_minimal()
plot9 <- ggplotly(g9)
plot9
```

The density plot highlights distinct differences in weight distribution between genders. Females generally exhibit lower weights, with a peak around 70 units, whereas males show peaks at 85 and 115 units, reflecting a tendency toward higher weights. An overlapping region between 80 and 90 units indicates common weight ranges for both genders, though the distinct peaks underscore gender-based differences. Weight in the dataset ranges from 39 to 173 units, with an average of 86.6 units, a median of 83 units, and a standard deviation of 26.2, indicating moderate variability.


Ridgeline Plot of Weight by Obesity Level.

```{r, message=FALSE, warning=FALSE}
ggplot(dataset, aes(x = weight, y = obesity_lev, fill = obesity_lev)) +
  geom_density_ridges(scale = 0.9, alpha = 0.6) +
  labs(title = "Ridgeline Plot of Weight by Obesity Level", x = "Weight", y = "Obesity Level") +
  theme_minimal() +
  theme(legend.position = "none")
# can't seem to make the interractive plot work 
```
The ridgeline plot further illustrates the relationship between weight and obesity levels. As obesity levels increase, the weight distribution shifts consistently toward higher values. Categories such as "Insufficient Weight" and "Normal Weight" cluster at lower ranges, while higher obesity types (I, II, and III) peak at significantly greater weights. This clear progression confirms a strong positive association between weight and obesity levels, reinforcing the centrality of weight in obesity classification. The dataset's average weight remains at 86.6 units with a standard deviation of 26.6, capturing the variability across different obesity categories.

##### Height and Weight

Researching deeply the relationship between height and weight, trends across obesity levels were examined using scatter plots, providing critical insights that reinforce theoretical expectations while contextualizing weight variations within height ranges for different obesity classifications.

Scatter Plot (height vs weight), colored by obesity level

```{r, message=FALSE, warning=FALSE}
g11 <- ggplot(dataset, aes(x = height, y = weight, color = obesity_lev)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, aes(group = obesity_lev)) +  # Adds a trend line for each obesity level
  ggtitle("Scatter Plot of Weight vs Height by Obesity Level") +
  theme_minimal() +
  labs(x = "Height", y = "Weight", color = "Obesity Level")
plot11 <- ggplotly(g11)
plot11
```

Facet Grid for Height and Weight by Obesity Level

```{r}
g12 <- ggplot(dataset, aes(x = height, y = weight)) +
  geom_point(alpha = 0.7, aes(color = obesity_lev)) +
  facet_wrap(~ obesity_lev) +
  ggtitle("Facet Grid of Weight and Height by Obesity Level") +
  theme_minimal() +
  labs(x = "Height", y = "Weight", color = "Obesity Level") +
  theme(legend.position = "none")
plot12 <- ggplotly(g12)
plot12

```

The first scatter plot presents an overview, illustrating the general trend of increasing weight with height, stratified by obesity levels. To better isolate and visualize these individual trends, the initial graph is expanded into a facet grid, offering a clearer perspective on the separate trends within each obesity category and highlighting distinct relationships and ranges.

Correlation between height and weight

```{r}
correlation_height_weight <- cor(dataset$height, dataset$weight, use = "complete.obs")
correlation_height_weight
```
The observed correlation between height and weight (r = 0.463) aligns with findings in existing literature, confirming a moderate positive relationship and reinforcing the expectation that taller individuals generally weigh more, though the strength of this association varies slightly across obesity levels.


With the analysis of age, height, and weight completed, attention shifts to exploring the remaining variables in the dataset. These variables, while less directly correlated with the target outcomes, offer critical insights into behavioral, lifestyle, and environmental factors that may influence obesity levels. 

##### Food between meals

```{r}

# Dodged Bar Chart for food_btw_meals by obesity levels
g13 <- ggplot(dataset, aes(x = food_btw_meals, fill = obesity_lev)) +
   geom_bar(position = "dodge", color = "black") +
   ggtitle("Dodged Bar Chart for Food Between Meals by Obesity Levels") +
   labs(x = "Food Between Meals", y = "Count", fill = "Obesity Levels") +
   theme_minimal() +
   theme(
         plot.title = element_text(hjust = 0.5, size = 14))

plot13 <- ggplotly(g13)
plot13

# Stacked Bar Chart of Food Between Meals by Obesity Level (Proportions within each Obesity Level)
g14 <- ggplot(dataset, aes(x = obesity_lev, fill = food_btw_meals)) +
    geom_bar(position = "fill") + # Stacked bar chart with proportions
    scale_y_continuous(labels = scales::percent_format(accuracy = 1)) + # Format y-axis as percentages
    ggtitle("Proportion of Food Between Meals Across Obesity Levels") + # Shortened and clear title
    labs(x = "Obesity Levels", y = "Proportion (%)", fill = "Food Between Meals") + # Correct axis and legend labels
    theme_minimal() +
    theme(
        axis.text.x = element_text(angle = 45, hjust = 1), # Rotate x-axis text for readability
        plot.title = element_text(hjust = 0.5, size = 14) # Center and style the title
    )
plot14 <- ggplotly(g14)
plot14
```

The charts provide a clear illustration of how the frequency of eating between meals varies across obesity levels. The most dominant behavior across all categories is "Sometimes," which peaks in intermediate levels like Normal Weight and Overweight Level I, reflecting a common pattern of moderate snacking. However, as obesity levels increase to Obesity Types I–III, the responses for "Frequently" and "Always" diminish, while "Sometimes" becomes even more prevalent. This shift could indicate that higher obesity levels are more associated with habitual moderate snacking rather than excessive meal-snacking frequency. On the other hand, "No" responses remain negligible across all obesity levels, suggesting that eating between meals is almost universal in this population. This pattern underscores the importance of examining not just the frequency but also the quality and context of snacking as potential contributors to obesity progression.

##### High-caloric food consumption


```{r}
# Grouped Bar Chart of High-Caloric Food by Obesity Level (Counts)
g16 <- ggplot(dataset, aes(x = obesity_lev, fill = caloric_food)) +
  geom_bar(
    position = "dodge",
    color = "black"
  ) +
  ggtitle("Grouped Bar Chart of High-Caloric Food Consumption Across Obesity Levels") +
  labs(x = "Obesity Levels", y = "Count", fill = "High-Caloric Food Consumption") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5, size = 14)
  )
plot16 <- ggplotly(g16)
plot16

```

The grouped bar chart highlights a clear trend of increased high-caloric food consumption as obesity levels rise. High-caloric food consumption ("yes") dominates across all categories, surpassing 75% of responses and becoming nearly universal in higher obesity levels (Obesity Type I–III). In contrast, "no" responses are more prominent in lower categories like Insufficient Weight and Normal Weight but remain relatively infrequent.


```{r}
percentage_high_caloric_consumers <- mean(dataset$caloric_food == "yes") * 100
percentage_high_caloric_consumers
```

More precisely, a notable 88.4% of participants report frequent consumption of high-calorie foods, a behavior strongly associated with weight gain. This underscores the critical importance of dietary interventions aimed at reducing high-calorie intake to address obesity progression effectively.

##### Alcohol consumption

Frequence in consumption of alcohol

```{r}
# Filter out "Always" responses from the dataset
filtered_dataset <- dataset %>%
  filter(freq_alcohol != "Always")

# Dodged Bar Chart for freq_alcohol by Obesity Levels (excluding "Always")
g17 <- ggplot(filtered_dataset, aes(x = freq_alcohol, fill = obesity_lev)) +
   geom_bar(position = "dodge", color = "black") +
   ggtitle("Dodged Bar Chart for Alcohol Consumption by Obesity Levels") +
   labs(x = "Alcohol Consumption Frequency", y = "Count", fill = "Obesity Levels") +
   theme_minimal() +
   theme(
         plot.title = element_text(hjust = 0.5, size = 14)) # Center and style the title

plot17 <- ggplotly(g17)
plot17


```

Regarding alcohol consumption, the chart shows that "Sometimes" is the dominant alcohol consumption frequency across all obesity levels, particularly in Normal Weight, Overweight Level I, and II categories. As obesity increases, "Frequently" becomes slightly more prominent, especially in Obesity Type III, while "No" responses decrease, being more common in lower obesity levels such as Insufficient and Normal Weight. The "Always" responses are excluded from this chart due to their near absence in the dataset, highlighting that excessive alcohol consumption is rare. This trend underlines the potential relationship between moderate-to-frequent alcohol consumption and higher obesity levels, emphasizing its importance for obesity-related behavioral research.

```{r}
# Prepare the data summary for 'Sometimes' and 'No' responses
data_summary <- dataset %>%
  filter(freq_alcohol %in% c("Sometimes", "No")) %>%
  group_by(obesity_lev, freq_alcohol) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(obesity_lev) %>%
  mutate(
    total = sum(count),
    proportion = count / total
  ) %>%
  ungroup()

# Visualization with updated title
g18 <- ggplot(data_summary, aes(x = obesity_lev, y = proportion, group = freq_alcohol, color = freq_alcohol)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +  # Format y-axis as percentages
  ggtitle("Proportion of 'Sometimes' and 'No' Alcohol Responses by Obesity Level") +
  labs(x = "Obesity Level", y = "Proportion (%)", color = "Alcohol Frequency") +
  scale_color_manual(values = c("No" = "purple", "Sometimes" = "gold")) + # Improved color scheme
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5, size = 14),  # Center and style title
    legend.position = "top"
  )
plot18 <- ggplotly(g18)
plot18
```

To better illustrate the trends in alcohol consumption frequency across obesity levels, this graph was created to highlight the shifting proportions of individuals consuming alcohol "Sometimes" and abstaining ("No"). The proportion of individuals who drink alcohol "Sometimes" shows a steady increase with higher obesity levels, peaking in Obesity_Type_III. Conversely, the proportion of those who abstain from alcohol decreases as obesity levels rise, suggesting an inverse relationship between abstention and obesity severity.

This pattern raises questions about the potential interaction between alcohol consumption frequency and caloric food preferences, as both behaviors appear to be associated with higher obesity levels. Investigating this interaction could provide insights into whether a combination of moderate alcohol consumption and high-calorie food preferences exerts a compounded effect on obesity risk. Understanding these combined lifestyle factors could inform strategies aimed at mitigating obesity progression more effectively.

##### Daily Calorie Monitoring

```{r}


# Dodged Bar Chart for calorie_check by Obesity Levels
g19 <- ggplot(dataset, aes(x = calorie_check, fill = obesity_lev)) +
   geom_bar(position = "dodge", color = "black") +
   ggtitle("    Dodged Bar Chart for the check of the calories by Obesity Levels") +
   labs(x = "High-Caloric Food Consumption", y = "Count", fill = "Obesity Levels") +
   theme_minimal() +
   theme(
         plot.title = element_text(hjust = 0.5, size = 14)) # Center and style the title
plot19 <- ggplotly(g19)
plot19

data_summary <- dataset %>%
  group_by(obesity_lev, calorie_check) %>%
  summarise(count = n(), .groups = "drop") %>%
  mutate(total = sum(count), proportion = count / total)

# Proportion of Calorie Checking by Obesity Level
g20 <- ggplot(data_summary, aes(x = obesity_lev, y = proportion, group = calorie_check, color = calorie_check)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  scale_y_continuous(labels = scales::percent) +
  scale_color_manual(values = c("no" = "lightcoral", "yes" = "lightblue")) +
  labs(title = "Proportion of Calorie Checking by Obesity Level", x = "Obesity Level", y = "Proportion", color = "Calorie Check") +
  theme_minimal() +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))
plot20 <- ggplotly(g20)
plot20

```

The Dodged Bar Chart highlights two main trends regarding calorie-checking behavior across obesity levels: a significant increase in "Yes" responses as obesity levels rise, particularly from Overweight Level II onward, and a decrease in "No" responses, which are more prevalent in lower obesity levels like Normal Weight and Insufficient Weight. The proportion graph simplifies these trends by clearly illustrating the proportional shift between "Yes" and "No" responses, making the contrast between lower and higher obesity levels more visually apparent. Together, these visualizations emphasize a potential association between obesity severity and an increased tendency to check calorie intake, suggesting heightened dietary awareness in higher obesity categories.

##### Vegetable consumption

```{r}
g21 <- ggplot(dataset, aes(x = vegetable_food)) +
  geom_histogram(aes(y =after_stat(density)), bins = 30, fill = "lightgreen", color = "black", alpha = 0.6) +
  geom_density(color = "darkgreen", linewidth = 1) +
  ggtitle("Histogram and Density of Vegetable Food Consumption") +
  theme_minimal() +
  labs(x = "Vegetable Food Consumption", y = "Density")
plot21 <- ggplotly(g21)
plot21

```

```{r, message=FALSE, warning=FALSE}
g22 <- ggplot(dataset, aes(x = weight, y = vegetable_food, color = obesity_lev)) +
    geom_point(alpha = 0.6) +
    geom_smooth(method = "loess", se = FALSE, color = "black") +
    labs(title = "Scatterplot of Weight vs Vegetable Food Consumption", 
         x = "Weight", 
         y = "Vegetable Food Consumption") +
    theme_minimal() +
    coord_cartesian(xlim= c(40, 135), ylim= c(2, 3))
plot22 <- ggplotly(g22)
plot22
```

The scatterplot provided with the trend line illustrates a distinct, non-linear relationship: vegetable consumption initially decreases as weight increases but then begins to rise again at higher weight levels.

This pattern suggests that individuals with lower weight, particularly those in the Insufficient Weight and Normal Weight categories, tend to report higher vegetable consumption. As weight progresses toward the Overweight categories, vegetable consumption decreases slightly, indicating a possible reduction in healthy dietary habits. However, at the upper end of the weight spectrum, corresponding to Obesity Type II and Obesity Type III, vegetable consumption increases again, potentially due to dietary interventions or awareness in this group.

The trend reveals two possible key insights:

-   A dip in vegetable consumption occurs in intermediate weight ranges, aligning with the overweight population.
-   The sharp increase in vegetable consumption among the most obese individuals may reflect lifestyle adjustments prompted by health concerns or medical advice.

##### Physical activity

```{r, warning=FALSE}
g23 <- ggplot(dataset, aes(x = physical_act)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "skyblue", color = "black", alpha = 0.6) +
  geom_density(color = "darkblue", linewidth = 1) +
  ggtitle("Histogram and Density of Physical Activity") +
  theme_minimal() +
  labs(x = "Physical Activity", y = "Density")
plot23 <- ggplotly(g23)
plot23
```

The histogram and density plot reveal that physical activity levels have distinct peaks at 0, 1, 2, and 3, suggesting that these values are common reported levels. Intermediate values, likely due to synthetic data or SMOTE, are also present but less frequent.


Violin plot by category

```{r}
g24 <- ggplot(dataset, aes(x = obesity_lev, y = physical_act, fill = obesity_lev)) +  
  geom_violin(trim = FALSE, alpha = 0.6) +
  geom_boxplot(width = 0.1, color = "black", fill = "white") +
    ggtitle("Violin Plot of Physical Activity by Obesity Level") +
  labs(x = "Obesity Level", y = "Physical Activity") +
  theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))
plot24 <- ggplotly(g24)
plot24
```

Physical activity levels show a slight decline as obesity levels increase, particularly evident in the narrowing distributions and lower medians observed for Obesity Type II and Obesity Type III categories. In contrast, the Insufficient Weight and Normal Weight groups exhibit higher physical activity levels, as reflected by their broader and more symmetrical distributions.

The graph reveals a distinct trend: individuals in lower obesity categories engage in more physical activity compared to those in higher obesity categories. This trend suggests an inverse relationship between physical activity and obesity levels.

##### Water consumption

```{r, warning=FALSE}
g25 <- ggplot(dataset, aes(x = ch2o)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "skyblue", color = "black", alpha = 0.6) +
  geom_density(color = "darkblue", size = 1) +
  ggtitle("Histogram and Density of Comsumption of Water") +
  theme_minimal() +
  labs(x = "CH2O", y = "Density")
plot25 <- ggplotly(g25)
plot25
```

This histogram and density plot of daily water consumption (CH2O) shows a clear peak at 2 liters, indicating that most individuals consume around this amount. This aligns with scientific literature, which generally recommends an average daily water intake of about 2 liters for optimal health.


Trend Line of Weight vs Water Consumption

```{r, message=FALSE}
# Scatterplot with a LOESS trend line
g26 <- ggplot(dataset, aes(x = weight, y = ch2o, color = obesity_lev)) +
    geom_point(alpha = 0.6) +
    geom_smooth(method = "loess", se = FALSE, color = "black") +
    labs(title = "Scatterplot of Weight vs Water Consumption", x = "Weight", y = "Water Consumption (ch2o)") +
    theme_minimal() +
coord_cartesian(xlim= c(35, 135))
plot26 <- ggplotly(g26)
plot26

```

The scatterplot visualizes the relationship between weight and water consumption (ch2o), categorized by obesity levels. The trend line reveals a slightly increasing pattern of water consumption as weight increases, though the relationship is relatively weak and mostly linear.

This pattern suggests that individuals with Insufficient Weight and Normal Weight categories generally report slightly lower water consumption compared to individuals in the higher weight categories, such as Obesity Type II and III. The increase in water consumption among higher weight groups could indicate attempts to adopt healthier habits or increased hydration needs due to larger body sizes. However, the relatively flat trend across most weight ranges suggests that water consumption does not vary dramatically across different weight categories, highlighting a potential area for targeted interventions to promote hydration as a component of healthy dietary behavior.

##### Technology utilization

Density of Use of Technology by Obesity Level

```{r}
g28 <- ggplot(dataset, aes(x = use_tech, fill = obesity_lev)) +
  geom_density(alpha = 0.5) +
  labs(title = "Density of Use of Technology by Obesity Level", x = "Use of Technology", y = "Density") +
  theme_minimal()
plot28 <- ggplotly(g28)
plot28
```

This density plot provides a perspective on the use of technology across different obesity levels. A striking feature is the sharp, dominant peak in Obesity Type III (yellow) around the value of 1. This pattern diverges notably from the smoother and more evenly distributed curves seen in other obesity categories, suggesting a unique behavioral trend in this group.

The peak indicates a strong clustering of individuals in Obesity Type III who report moderate use of technology, which may reflect consistent engagement with technology-based activities such as sedentary work, entertainment, or even health-monitoring applications. In contrast, other obesity categories, such as Obesity Type II and Overweight Level II, exhibit more balanced distributions without a single dominant peak, hinting at more varied technology usage patterns.

This observation raises interesting questions about the role of technology in shaping lifestyle behaviors in Obesity Type III individuals. It may point to a reliance on technology that correlates with a sedentary lifestyle, a known risk factor for obesity. Alternatively, it could reflect targeted interventions or habits specific to this group.


The Exploratory Data Analysis (EDA) phase provided a comprehensive understanding of the dataset, offering key insights into the relationships between various behavioral, lifestyle, and demographic factors and obesity levels. By focusing on critical variables the EDA revealed patterns and trends that are integral to the modeling process.



